---
title: "03 Build Wave 11 Metadata"
author: "Yiwen Xiang"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# 03 Build Wave 11 Metadata

This script builds the labelled Wave 11 variable metadata and value-code dictionaries used for variable matching.
If the imported Wave 11 RDS file is missing, it runs the Wave 11 import step first.

## Load imported data and helper functions

```{r}
source(here::here("script", "00_packages.R"))
source(here::here("script", "02_functions.R"))


wave11_datasets_file <- file.path(processed_path, "01_wave11_datasets.rds")

if (file.exists(wave11_datasets_file)) {
  w11_datasets <- readRDS(wave11_datasets_file)
} else {
  message("Processed Wave 11 dataset RDS not found. Running 01_import_wave11.Rmd.")
  # Use a temporary extracted R script so the submission folder keeps only the Rmd version.
  source(knitr::purl(here::here("script", "01_import_wave11.Rmd"), output = tempfile(fileext = ".R"), quiet = TRUE))
}
```

## List dictionary files

```{r}
dict_files <- tibble::tribble(
  ~table, ~file_name,
  "w11_main", "5050_wave_11_elsa_data_eul_data_dictionary.xlsx",
  "w11_ifs", "5050_wave_11_ifs_derived_variables_ukda_data_dictionary.xlsx",
  "w11_financial", "5050_wave_11_financial_derived_variables_ukda_data_dictionary.xlsx",
  "w11_nurse", "5050_wave_11_elsa_nurse_data_eul_data_dictionary.xlsx",
  "w11_pensiongrid", "5050_wave_11_elsa_pensiongrid_eul_data_dictionary.xlsx",
  "w11_lifehistory", "5050_wave_11_elsa_lifehistory02_data_dictionary.xlsx",
  "w11_eol", "5050_w11_elsa_eol_data_dictionary.xlsx"
)
```

## Check required objects

```{r}
required_objects <- c(
  "w11_datasets",
  "dict_files",
  "wave11_path",
  "processed_path"
)

missing_objects <- required_objects[!sapply(required_objects, exists)]

if (length(missing_objects) > 0) {
  stop(
    paste(
      "Missing required objects:",
      paste(missing_objects, collapse = ", ")
    )
  )
}

required_functions <- c(
  "build_variable_dictionary",
  "build_value_dictionary"
)

missing_functions <- required_functions[!sapply(required_functions, exists)]

if (length(missing_functions) > 0) {
  stop(
    paste(
      "Missing required functions:",
      paste(missing_functions, collapse = ", ")
    )
  )
}
```

## Build variable metadata

```{r}
var_w11_all_labeled <- build_variable_dictionary(
  w11_datasets = w11_datasets,
  dict_files = dict_files,
  wave11_path = wave11_path
)
```

## Build value-code dictionary

```{r}
value_dict_w11_all <- build_value_dictionary(
  dict_files = dict_files,
  wave11_path = wave11_path
)
```

## Check metadata coverage

```{r}
var_label_summary <- var_w11_all_labeled %>%
  dplyr::group_by(table) %>%
  dplyr::summarise(
    n_variables = dplyr::n(),
    n_with_label = sum(!is.na(label)),
    n_without_label = sum(is.na(label)),
    pct_with_label = round(n_with_label / n_variables * 100, 2),
    .groups = "drop"
  )

var_label_summary
```

## Save metadata objects

```{r}
saveRDS(
  var_w11_all_labeled,
  file.path(processed_path, "03var_w11_all_labeled.rds")
)

saveRDS(
  value_dict_w11_all,
  file.path(processed_path, "03value_dict_w11_all.rds")
)

saveRDS(
  var_label_summary,
  file.path(processed_path, "03var_label_summary.rds")
)

message("Wave 11 metadata has been built successfully.")
message("Saved: data/processed/03var_w11_all_labeled.rds")
message("Saved: data/processed/03value_dict_w11_all.rds")
message("Saved: data/processed/03var_label_summary.rds")
```

