---
title: "05 Filter Analysis Sample"
author: "Yiwen Xiang"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: true
---

# 05 Filter Analysis Sample

This file filters the Wave 11 base analytical sample before constructing the CASP-19 outcome.
The main restrictions are Wave 11 core membership, valid age and valid sex.

## Load inputs

```{r, include=FALSE}
source(here::here("script", "00_packages.R"))
source(here::here("script", "02_functions.R"))

w11_datasets <- readRDS(file.path(processed_path, "01_wave11_datasets.rds"))
data11_main <- w11_datasets$w11_main

dataset_summary <- readRDS(file.path(processed_path, "01_wave11_dataset_summary.rds"))
var_w11_all_labeled <- readRDS(file.path(processed_path, "03var_w11_all_labeled.rds"))
value_dict_w11_all <- readRDS(file.path(processed_path, "03value_dict_w11_all.rds"))
var_label_summary <- readRDS(file.path(processed_path, "03var_label_summary.rds"))
```

## Check imported data

```{r}
dataset_summary
```

## Check respondent ID and core member status

```{r}
id_check <- data11_main %>%
  summarise(
    n_rows = n(),
    n_unique_id = n_distinct(idauniq),
    n_missing_id = sum(is.na(idauniq)),
    duplicated_id_count = n_rows - n_unique_id
  )

id_check
```

```{r}
corepartner_check <- data11_main %>%
  count(corepartner, name = "n", sort = TRUE) %>%
  mutate(
    percentage = round(n / sum(n) * 100, 2)
  )

corepartner_check
```
```{r}
search_code(
  var_names = "corepartner",
  value_dict = value_dict_w11_all,
  table_name = "w11_main"
)
```

According to the Wave 11 data dictionary, `corepartner` identifies whether a respondent is a core member or a partner: `0` represents Partner, and `1` represents Core Member. Since the analysis focuses on eligible core Wave 11 respondents, only records with `corepartner == 1` were retained in the analytical sample.


```{r}
data11_main %>%
  summarise(
    n_total = n(),
    n_corepartner_missing = sum(is.na(corepartner)),
    pct_corepartner_missing = round(mean(is.na(corepartner)) * 100, 2)
  )
```

## Restrict to core members

```{r}
data11_core <- data11_main %>%
  filter(corepartner == 1)

nrow(data11_core)
```

## Age variable check and cleaning

### Confirm age variable

```{r}
var_w11_all_labeled %>%
  filter(table == "w11_main", variable == "indager") %>%
  select(table, variable, label, type, n_missing)
```

```{r}
search_code(
  var_names = "indager",
  value_dict = value_dict_w11_all,
  table_name = "w11_main"
)
```

### Raw age distribution

```{r}
summary(data11_core$indager)
```

```{r}
data11_core %>%
  count(indager, name = "n", sort = FALSE) %>%
  arrange(indager)
```

```{r}
age_special_code_check <- data11_core %>%
  filter(indager < 0) %>%
  count(indager, corepartner, name = "n", sort = TRUE)

age_special_code_check
```

### Clean age variable

```{r}
# Keep -7 as 90 because ELSA collapses ages 90+ for disclosure control.
# Other negative age codes are treated as missing.

data11_core_age <- data11_core %>%
  mutate(
    age = case_when(
      indager == -7 ~ 90,
      indager %in% c(-9, -8, -1) ~ NA_real_,
      indager >= 0 ~ as.numeric(indager),
      TRUE ~ NA_real_
    ),
    age_group_90plus = case_when(
      indager == -7 ~ "90+",
      indager >= 0 ~ as.character(indager),
      TRUE ~ NA_character_
    )
  )
```


## Sex variable check and cleaning

### Confirm sex variable

```{r}
var_w11_all_labeled %>%
  filter(table == "w11_main", variable == "indsex") %>%
  select(table, variable, label, type, n_missing)
```

```{r}
search_code(
  var_names = "indsex",
  value_dict = value_dict_w11_all,
  table_name = "w11_main"
)
```

### Raw sex distribution

```{r}
data11_core_age %>%
  count(indsex, name = "n", sort = FALSE) %>%
  mutate(
    percentage = round(n / sum(n) * 100, 2)
  )
```

### Clean sex variable

```{r}
# Negative sex codes are non-substantive responses and are treated as missing.

data11_core_age_sex <- data11_core_age %>%
  mutate(
    sex = case_when(
      indsex == 1 ~ "Male",
      indsex == 2 ~ "Female",
      indsex %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    sex = factor(sex, levels = c("Male", "Female"))
  )
```

```{r}
sex_distribution_before_filter <- data11_core_age_sex %>%
  count(sex, name = "n") %>%
  mutate(
    percentage = round(n / sum(n) * 100, 2)
  )

sex_distribution_before_filter
```

```{r}
data11_core_age_sex %>%
  summarise(
    n_total = n(),
    n_sex_missing = sum(is.na(sex)),
    pct_sex_missing = round(mean(is.na(sex)) * 100, 2)
  )
```

## Build initial analysis sample

```{r}
# This base sample is used for outcome calculation and later modelling steps.

data11_analysis_base <- data11_core_age_sex %>%
  filter(
    !is.na(age),
    age >= 50,
    !is.na(sex)
  )

data11_analysis_base
```

```{r}
nrow(data11_analysis_base)
```


## Overview of final base analysis sample

### Age summary

```{r}
age_summary_final <- data11_analysis_base %>%
  summarise(
    n = n(),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    q1_age = quantile(age, 0.25, na.rm = TRUE),
    q3_age = quantile(age, 0.75, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE),
    n_age_90plus = sum(indager == -7, na.rm = TRUE),
    pct_age_90plus = round(mean(indager == -7, na.rm = TRUE) * 100, 2)
  )

age_summary_final
```

```{r}
# Age group distribution

age_group_summary_final <- data11_analysis_base %>%
  mutate(
    age_group = case_when(
      age >= 50 & age < 60 ~ "50-59",
      age >= 60 & age < 70 ~ "60-69",
      age >= 70 & age < 80 ~ "70-79",
      age >= 80 & age < 90 ~ "80-89",
      age >= 90 ~ "90+",
      TRUE ~ NA_character_
    ),
    age_group = factor(
      age_group,
      levels = c("50-59", "60-69", "70-79", "80-89", "90+")
    )
  ) %>%
  count(age_group, name = "n") %>%
  mutate(
    percentage = round(n / sum(n) * 100, 2)
  )

age_group_summary_final
```

### Sex summary

```{r}
sex_summary_final <- data11_analysis_base %>%
  count(sex, name = "n") %>%
  mutate(
    percentage = round(n / sum(n) * 100, 2)
  )

sex_summary_final
```

### Age by sex summary

```{r}
age_by_sex_summary <- data11_analysis_base %>%
  group_by(sex) %>%
  summarise(
    n = n(),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE),
    n_age_90plus = sum(indager == -7, na.rm = TRUE),
    .groups = "drop"
  )

age_by_sex_summary
```

## Plots for base analysis sample

### Age distribution

```{r}
mean_age <- mean(data11_analysis_base$age, na.rm = TRUE)

ggplot(data11_analysis_base, aes(x = age)) +
  geom_histogram(
    binwidth = 5,
    boundary = 50,
    closed = "left",
    fill = "#74ADD1",
    color = "white"
  ) +
  geom_vline(
    xintercept = mean_age,
    color = "red",
    linewidth = 1,
    linetype = "dashed"
  ) +
  scale_x_continuous(
    breaks = seq(50, 90, by = 5),
    labels = c(seq(50, 85, by = 5), "90+")
  ) +
  labs(
    title = "Age distribution in the base analysis sample",
    subtitle = paste0(
      "Dashed red line = mean age: ",
      round(mean_age, 1),
      "; age 90+ is top-coded as 90"
    ),
    x = "Age",
    y = "Number of participants"
  ) +
  theme_minimal(base_size = 13)
```

### Sex distribution

```{r}
sex_plot_data <- data11_analysis_base %>%
  count(sex, name = "n") %>%
  mutate(
    percentage = n / sum(n) * 100,
    label = paste0(n, " (", round(percentage, 1), "%)")
  )

ggplot(sex_plot_data, aes(x = sex, y = n, fill = sex)) +
  geom_col(width = 0.6, color = "white") +
  geom_text(
    aes(label = label),
    vjust = -0.4,
    size = 4
  ) +
  scale_fill_manual(
    values = c(
      "Male" = "#74ADD1",
      "Female" = "#F46D43"
    )
  ) +
  labs(
    title = "Sex distribution in the base analysis sample",
    x = "Sex",
    y = "Number of participants",
    fill = "Sex"
  ) +
  theme_minimal(base_size = 13)
```

### Age distribution by sex

```{r}
ggplot(data11_analysis_base, aes(x = sex, y = age, fill = sex)) +
  geom_boxplot(width = 0.6, outlier.alpha = 0.4) +
  scale_fill_manual(
    values = c(
      "Male" = "#74ADD1",
      "Female" = "#F46D43"
    )
  ) +
  scale_y_continuous(
    breaks = seq(50, 90, by = 5),
    labels = c(seq(50, 85, by = 5), "90+")
  ) +
  labs(
    title = "Age by sex in the base analysis sample",
    subtitle = "Age 90+ is top-coded as 90",
    x = "Sex",
    y = "Age",
    fill = "Sex"
  ) +
  theme_minimal(base_size = 13)
```

## Save base analysis sample

```{r}
# Save the filtered base analysis sample.
# This file will be used in 06_outcome_casp19.Rmd and later analysis files.

saveRDS(
  data11_analysis_base,
  file.path(processed_path, "05_wave11_analysis_sample.rds")
)

```
