---
title: "06 CASP-19 Outcome"
author: "Yiwen Xiang"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: true
---

# 06 CASP-19 Outcome

This file constructs the CASP-19 total score and the binary low-QoL outcome. The cut-off was based on the distribution of CASP-19 total scores in the analytical sample: the lower quartile corresponded to a score of 35. Therefore, respondents with a CASP-19 total score <= 35 were classified as low QoL, while those with scores > 35 were classified as non-low QoL.

## Load inputs

```{r, include=FALSE}
source(here::here("script", "00_packages.R"))
source(here::here("script", "02_functions.R"))

w11_datasets <- readRDS(file.path(processed_path, "01_wave11_datasets.rds"))
data11_main <- w11_datasets$w11_main

dataset_summary <- readRDS(file.path(processed_path, "01_wave11_dataset_summary.rds"))
var_w11_all_labeled <- readRDS(file.path(processed_path, "03var_w11_all_labeled.rds"))
value_dict_w11_all <- readRDS(file.path(processed_path, "03value_dict_w11_all.rds"))
var_label_summary <- readRDS(file.path(processed_path, "03var_label_summary.rds"))

picture_path <- here::here("pictures")
```

## Check imported data and metadata

```{r}
dataset_summary
```

```{r}
head(var_w11_all_labeled)
```

```{r}
var_label_summary
```


## Search CASP-19 / QoL outcome variables

### Search CASP-19 item variables

```{r}
casp_general_search <- search_vars(
  keywords = c(
    "CASP",
    "CASP-19",
    "CASP19",
    "quality of life",
    "QoL",
    "control autonomy self-realisation pleasure",
    "control autonomy self-realization pleasure"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% names(w11_datasets))

casp_general_search
```

### Search existing CASP-19 total score variable

```{r}
casp_score_search <- search_vars(
  keywords = c(
    "CASP score",
    "CASP total",
    "CASP total score",
    "CASP-19 score",
    "CASP19 score",
    "CASP-19 total",
    "CASP19 total",
    "quality of life score",
    "QoL score",
    "QoL total",
    "total quality of life",
    "summary CASP",
    "summed CASP",
    "CASP scale",
    "CASP index",
    "life satisfaction score",
    "wellbeing score",
    "well-being score"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% names(w11_datasets))

casp_score_search
```

No directly usable CASP-19 total score variable was confirmed in the searched
metadata. Therefore, the total score is calculated from the 19 CASP items.

### Load the analytical sample

```{r}
data11_analysis_base <- readRDS(
  file.path(processed_path, "05_wave11_analysis_sample.rds")
)

data_casp <- data11_analysis_base

nrow(data_casp)
```


## Confirm CASP-19 item variables

```{r}
# CASP-19 items are expected to be scqola to scqols.
casp_items <- paste0("scqol", letters[1:19])

casp_items
```

```{r}
casp_items <- paste0("scqol", letters[1:19])

negative_items <- c(
  "scqola",
  "scqolb",
  "scqold",
  "scqolf",
  "scqolh",
  "scqoli"
)

positive_items <- setdiff(casp_items, negative_items)

casp_raw_items <- paste0(casp_items, "_raw")
casp_score_items <- paste0(casp_items, "_score")

casp_items
negative_items
positive_items
```

```{r}
missing_casp_items <- setdiff(casp_items, names(data_casp))

if (length(missing_casp_items) > 0) {
  stop(
    "These CASP-19 variables are missing from data_casp: ",
    paste(missing_casp_items, collapse = ", ")
  )
}
```

## Check raw CASP-19 missingness and invalid codes

```{r}
casp_missing_raw <- purrr::map_dfr(casp_items, function(v) {
  x <- suppressWarnings(as.numeric(data_casp[[v]]))
  
  tibble::tibble(
    variable = v,
    n_total = length(x),
    n_raw_na = sum(is.na(x)),
    n_negative_code = sum(!is.na(x) & x < 0),
    n_outside_1_4_positive = sum(!is.na(x) & x >= 0 & !x %in% 1:4),
    pct_negative_code = round(mean(!is.na(x) & x < 0) * 100, 2)
  )
})

casp_missing_raw
```

## Create cleaned raw CASP item variables

```{r}
clean_casp_item <- function(x) {
  x <- suppressWarnings(as.numeric(x))
  
  # ELSA negative response codes are treated as missing
  x[x < 0] <- NA_real_
  
  # Valid raw CASP item responses should only be 1, 2, 3, or 4
  x[!is.na(x) & !x %in% 1:4] <- NA_real_
  
  x
}

data_casp <- data_casp %>%
  mutate(
    across(
      all_of(casp_items),
      clean_casp_item,
      .names = "{.col}_raw"
    )
  )
```

```{r}
casp_clean_missing_by_item <- purrr::map_dfr(casp_raw_items, function(v) {
  x <- data_casp[[v]]
  
  tibble::tibble(
    variable = v,
    n_missing = sum(is.na(x)),
    pct_missing = round(mean(is.na(x)) * 100, 2),
    n_valid = sum(!is.na(x)),
    min_value = min(x, na.rm = TRUE),
    max_value = max(x, na.rm = TRUE)
  )
})

casp_clean_missing_by_item
```


## Count valid and missing CASP-19 items per participant
```{r}
data_casp <- data_casp %>%
  mutate(
    casp19_n_valid_items = rowSums(!is.na(select(., all_of(casp_raw_items)))),
    casp19_n_missing_items = length(casp_raw_items) - casp19_n_valid_items
  )

table(data_casp$casp19_n_missing_items, useNA = "ifany")
```

```{r}
data_casp %>%
  summarise(
    n_total = n(),
    n_complete_19_items = sum(casp19_n_valid_items == 19),
    n_incomplete_items = sum(casp19_n_valid_items < 19),
    pct_complete_19_items = round(mean(casp19_n_valid_items == 19) * 100, 2)
  )
```
This check shows that 5,314 of the 6,422 core Wave 11 respondents had complete CASP-19 item data. Since this study used a complete-case approach for the outcome variable, the 1,108 respondents with missing or invalid CASP-19 item responses were excluded from the outcome-valid analytical sample.


## Recode CASP-19 items into 0-3 score variables
Before calculating the CASP-19 total score, all items were recoded onto a 0-3 scale so that higher values consistently represented better QoL. Negatively worded items were scored as `raw response - 1`, while positively worded items were reverse-coded as `4 - raw response`. The 19 recoded item scores were then summed to create the CASP-19 total score, with a possible range from 0 to 57.
```{r}
for (v in negative_items) {
  raw_var <- paste0(v, "_raw")
  score_var <- paste0(v, "_score")
  
  data_casp[[score_var]] <- data_casp[[raw_var]] - 1
}

for (v in positive_items) {
  raw_var <- paste0(v, "_raw")
  score_var <- paste0(v, "_score")
  
  data_casp[[score_var]] <- 4 - data_casp[[raw_var]]
}
```

```{r}
casp_scored_range_check <- purrr::map_dfr(casp_score_items, function(v) {
  x <- data_casp[[v]]
  
  tibble::tibble(
    variable = v,
    min_value = min(x, na.rm = TRUE),
    max_value = max(x, na.rm = TRUE),
    n_outside_0_3 = sum(!is.na(x) & !x %in% 0:3)
  )
})

casp_scored_range_check
```

```{r}
if (any(casp_scored_range_check$n_outside_0_3 > 0)) {
  stop("Some scored CASP items are outside the expected range 0-3.")
}
```

## Calculate CASP-19 total score
```{r}
data_casp <- data_casp %>%
  mutate(
    casp19_total = ifelse(
      casp19_n_missing_items == 0,
      rowSums(select(., all_of(casp_score_items)), na.rm = FALSE),
      NA_real_
    )
  )
```

```{r}
casp19_range_check <- data_casp %>%
  summarise(
    n_valid = sum(!is.na(casp19_total)),
    n_missing = sum(is.na(casp19_total)),
    min_score = min(casp19_total, na.rm = TRUE),
    max_score = max(casp19_total, na.rm = TRUE),
    n_below_0 = sum(casp19_total < 0, na.rm = TRUE),
    n_above_57 = sum(casp19_total > 57, na.rm = TRUE)
  )

casp19_range_check
```

## Descriptive statistics for QoL score
```{r}
casp19_summary <- data_casp %>%
  summarise(
    n_total = n(),
    n_valid = sum(!is.na(casp19_total)),
    n_missing = sum(is.na(casp19_total)),
    missing_pct = round(mean(is.na(casp19_total)) * 100, 2),
    mean = round(mean(casp19_total, na.rm = TRUE), 2),
    median = round(median(casp19_total, na.rm = TRUE), 2),
    sd = round(sd(casp19_total, na.rm = TRUE), 2),
    min = min(casp19_total, na.rm = TRUE),
    q1 = round(quantile(casp19_total, 0.25, na.rm = TRUE), 2),
    q3 = round(quantile(casp19_total, 0.75, na.rm = TRUE), 2),
    max = max(casp19_total, na.rm = TRUE)
  )

casp19_summary
```

## Define low QoL using the fixed CASP-19 cutoff
```{r}
qol_q1_cutoff <- 35

qol_q1_cutoff
```

```{r}
data_casp <- data_casp %>%
  mutate(
    qol_group = case_when(
      is.na(casp19_total) ~ NA_character_,
      casp19_total <= qol_q1_cutoff ~ "Low QoL",
      casp19_total > qol_q1_cutoff ~ "Non-low QoL"
    ),
    
    low_qol = case_when(
      is.na(casp19_total) ~ NA_real_,
      casp19_total <= qol_q1_cutoff ~ 1,
      casp19_total > qol_q1_cutoff ~ 0
    ),
    
    qol_group = factor(
      qol_group,
      levels = c("Non-low QoL", "Low QoL")
    )
  )
```

```{r}
qol_group_balance <- data_casp %>%
  filter(!is.na(qol_group)) %>%
  count(qol_group, name = "n") %>%
  mutate(
    pct = round(n / sum(n) * 100, 2)
  )

qol_group_balance
```

## Plot CASP-19 distribution


The optimized plot first restricts the figure to respondents with valid CASP-19 total scores, which matches the complete-case outcome analysis and avoids missing-value warnings. The display range is controlled with `coord_cartesian()` rather than scale limits, so boundary values are not removed. The mean and lower-quartile cut-off lines are drawn as fixed-height segments and labelled directly in the plot, making the figure cleaner and easier to interpret.
```{r}
data_casp_plot <- data_casp %>%
  filter(!is.na(casp19_total))

mean_casp19 <- mean(data_casp_plot$casp19_total)

y_line_top <- 490
y_label_top <- 515

casp19_distribution_plot <- ggplot(data_casp_plot, aes(x = casp19_total)) +
  geom_histogram(
    binwidth = 2,
    boundary = 0,
    closed = "left",
    fill = "#74ADD1",
    color = "white",
    linewidth = 0.25,
    alpha = 0.9
  ) +
  annotate(
    "segment",
    x = mean_casp19,
    xend = mean_casp19,
    y = 0,
    yend = y_line_top,
    color = "red",
    linewidth = 1,
    linetype = "dashed"
  ) +
  annotate(
    "segment",
    x = qol_q1_cutoff,
    xend = qol_q1_cutoff,
    y = 0,
    yend = y_line_top,
    color = "#2C7BB6",
    linewidth = 1,
    linetype = "dotted"
  ) +
  annotate(
    "text",
    x = mean_casp19,
    y = y_label_top,
    label = "Mean",
    color = "red",
    size = 3.5,
    hjust = 0.5
  ) +
  annotate(
    "text",
    x = qol_q1_cutoff,
    y = y_label_top,
    label = "Q1 cut-off = 35",
    color = "#2C7BB6",
    size = 3.5,
    hjust = 0.5
  ) +
  scale_x_continuous(
    breaks = seq(0, 55, by = 5),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(
    breaks = seq(0, 500, by = 100),
    expand = expansion(mult = c(0, 0))
  ) +
  coord_cartesian(
    xlim = c(0, 57),
    ylim = c(0, 540),
    clip = "off"
  ) +
  labs(
    title = "Distribution of CASP-19 total score",
    x = "CASP-19 total score",
    y = "Number of participants"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey88", linewidth = 0.4),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

casp19_distribution_plot
```

```{r}
ggsave(
  filename = file.path(picture_path, "06_casp19_total_score_distribution.jpg"),
  plot = casp19_distribution_plot,
  width = 8.5,
  height = 5.2,
  units = "in",
  dpi = 300,
  bg = "white"
)

```


### final output check
```{r}
list(
  n_total = nrow(data_casp),
  n_valid_casp19 = sum(!is.na(data_casp$casp19_total)),
  n_missing_casp19 = sum(is.na(data_casp$casp19_total)),
  casp19_min = min(data_casp$casp19_total, na.rm = TRUE),
  casp19_max = max(data_casp$casp19_total, na.rm = TRUE),
  q1_cutoff = qol_q1_cutoff,
  low_qol_n = sum(data_casp$low_qol == 1, na.rm = TRUE),
  non_low_qol_n = sum(data_casp$low_qol == 0, na.rm = TRUE)
)
```


## Create the outcome-valid analytical sample
```{r}
# Create outcome-valid analysis sample

data11_analysis_outcome_valid <- data_casp %>%
  filter(
    !is.na(casp19_total),
    casp19_total >= 0,
    casp19_total <= 57
  )

outcome_valid_summary <- tibble::tibble(
  n_base_sample = nrow(data_casp),
  n_outcome_valid = nrow(data11_analysis_outcome_valid),
  n_outcome_excluded = nrow(data_casp) - nrow(data11_analysis_outcome_valid),
  pct_outcome_excluded = round(
    (nrow(data_casp) - nrow(data11_analysis_outcome_valid)) / nrow(data_casp) * 100,
    2
  )
)

outcome_valid_summary

# Save outcome-valid full analysis sample

saveRDS(
  data11_analysis_outcome_valid,
  file.path(processed_path, "06_wave11_outcome.rds")
)
```


