---
title: "07_EDA01_base_features"
author: "Yiwen Xiang"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: true
---

This file builds the base Wave 11 modelling dataset from the outcome-valid analytical sample and the matched predictor fields in `Preliminary Feature Selection.xlsx`. Additional candidate features are searched and added in `07_EDA02_additional_features.Rmd`.

# 1. Setup

```{r, include=FALSE}
source(here::here("script", "00_packages.R"))

w11_datasets <- readRDS(file.path(processed_path, "01_wave11_datasets.rds"))
data11_main <- w11_datasets$w11_main
data11_ifs <- w11_datasets$w11_ifs
data11_financial <- w11_datasets$w11_financial
data11_nurse <- w11_datasets$w11_nurse
```



```{r}
data11_analysis_base <- readRDS(
  file.path(processed_path, "06_wave11_outcome.rds")
)

data11_analysis_outcome_valid <- data11_analysis_base

head(data11_analysis_base)
```


```{r}
dim(data11_analysis_base)
```

```{r}
selected_features <- readxl::read_excel(
  here::here("data", "upload", "Preliminary Feature Selection.xlsx")
)

selected_features
```

```{r}
feature_map <- selected_features %>%
  mutate(
    table = str_squish(as.character(table)),
    column_name = str_squish(as.character(column_name)),
    features = str_squish(as.character(features))
  ) %>%
  filter(!is.na(column_name), column_name != "")

feature_map %>%
  count(table)
```

```{r}
financial_cols_needed <- feature_map %>%
  filter(table == "w11_financial") %>%
  pull(column_name) %>%
  unique()

financial_cols_needed
```

```{r}
data11_financial_selected <- data11_financial %>%
  select(idauniq, all_of(financial_cols_needed))

data11_financial_selected
```

```{r}
ifs_cols_needed <- feature_map %>%
  filter(table == "w11_ifs") %>%
  pull(column_name) %>%
  unique()

ifs_cols_needed
```

```{r}
data11_ifs_selected <- data11_ifs %>%
  select(idauniq, all_of(ifs_cols_needed))

data11_ifs_selected
```

```{r}
nurse_cols_needed <- feature_map %>%
  filter(table == "w11_nurse") %>%
  pull(column_name) %>%
  unique()

nurse_cols_needed
```

```{r}
data11_nurse_selected <- data11_nurse %>%
  select(idauniq, all_of(nurse_cols_needed))

data11_nurse_selected
```

```{r}
data11_analysis_full <- data11_analysis_base %>%
  left_join(
    data11_ifs_selected,
    by = "idauniq",
    relationship = "one-to-one"
  ) %>%
  left_join(
    data11_financial_selected,
    by = "idauniq",
    relationship = "one-to-one"
  ) %>%
  left_join(
    data11_nurse_selected,
    by = "idauniq",
    relationship = "one-to-one"
  )
```

```{r}
nrow(data11_analysis_base)
nrow(data11_analysis_full)

stopifnot(nrow(data11_analysis_base) == nrow(data11_analysis_full))
```

```{r}
all_feature_cols <- feature_map %>%
  pull(column_name) %>%
  unique()

id_cols <- intersect(
  c("idauniq", "idahhw11", "perid","casp19_total"), 
  names(data11_analysis_full)
)

data11_selected_features <- data11_analysis_full %>%
  select(
    all_of(id_cols),
    all_of(all_feature_cols)
  )

data11_selected_features
```

```{r}
dim(data11_selected_features)

names(data11_selected_features)
```
At this stage, the outcome-valid analytical sample has been joined with the matched Wave 11 predictor fields from the main, IFS, financial and nurse datasets. The resulting dataset now contains the respondent identifiers, CASP-19 outcome variables, and the raw candidate feature columns needed for feature construction.

The next step is to recode these raw variables into analysis-ready predictors. The following sections derive demographic, socioeconomic, physical health, mental health, social participation, behavioural, and cognitive variables before saving the pre-screening modelling dataset.

# 3. Demographic variables

Age, sex, marital or partnership status, and education level were recoded into analysis-ready variables.
```{r}
data11_selected_features %>%
  count(indsex, sort = TRUE)
```
```{r}
data11_selected_features %>%
  summarise(
    n_total = n(),
    n_missing = sum(is.na(indsex)),
    missing_percent = round(mean(is.na(indsex)) * 100, 2)
  )
```


## 3.1 age
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    age = if_else(indager == -7, 90, indager)
  )
```

## 3.2 sex
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    sex = case_when(
      indsex == 1 ~ "Male",
      indsex == 2 ~ "Female",
      TRUE ~ NA_character_
    ),
    sex = factor(
      sex,
      levels = c("Male", "Female")
    )
  )
```

## 3.3 education level
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    education_level = case_when(
      fqendm %in% c(2, 3) ~ "Low education",
      fqendm %in% c(4, 5) ~ "Middle education",
      fqendm %in% c(6, 7) ~ "Post-secondary education",
      fqendm == 8 ~ "Higher education",
      fqendm %in% c(-9, -8, -1, 1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    education_level = factor(
      education_level,
      levels = c(
        "Low education",
        "Middle education",
        "Post-secondary education",
        "Higher education"
      ),
      ordered = TRUE
    )
  )
```

## 3.4 marital / partnership status
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    marital_status = case_when(
      dimarr %in% c(-9, -8, -1, -3) ~ NA_character_,
      couple == 2 ~ "Cohabiting",
      couple == 1 | dimarr %in% c(2, 3) ~ "Married/civil partnered",
      dimarr == 1 ~ "Never married",
      dimarr %in% c(4, 5) ~ "Separated/divorced",
      dimarr == 6 ~ "Widowed",
      TRUE ~ NA_character_
    ),
    marital_status = factor(
      marital_status,
      levels = c(
        "Married/civil partnered",
        "Cohabiting",
        "Never married",
        "Separated/divorced",
        "Widowed"
      )
    )
  )
```



# 4. Socioeconomic variables

Wealth, work status, housing tenure, and financial difficulty were recoded using grouped categories suitable for modelling.
## 4.1 wealth quintile
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    wealth_quintile = factor(
      totwq5_bu_s,
      levels = c(1, 2, 3, 4, 5),
      labels = c(
        "Lowest wealth quintile",
        "Second wealth quintile",
        "Middle wealth quintile",
        "Fourth wealth quintile",
        "Highest wealth quintile"
      ),
      ordered = TRUE
    )
  )
```

## 4.2 Work status
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    dhwork = case_when(
      dhwork == 1 ~ "Yes",
      dhwork == 2 ~ "No",
      dhwork == -1 ~ NA_character_,
      TRUE ~ NA_character_
    ),
    dhwork = factor(
      dhwork,
      levels = c("No", "Yes")
    )
  )
```

## 4.3 Housing tenure
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    housing_tenure = case_when(
      hotenu == 1 ~ "Own outright",
      hotenu %in% c(2, 3) ~ "Mortgage/shared ownership",
      hotenu == 4 ~ "Rent",
      hotenu %in% c(5, 6) ~ "Other housing arrangement",
      hotenu %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    housing_tenure = factor(
      housing_tenure,
      levels = c(
        "Own outright",
        "Mortgage/shared ownership",
        "Rent",
        "Other housing arrangement"
      )
    )
  )
```

## 4.4 Financial difficulty
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    financial_difficulty = case_when(
      iafcon %in% c(1, 2) ~ "Low financial difficulty",
      iafcon %in% c(3, 4) ~ "Moderate financial difficulty",
      iafcon %in% c(5, 6) ~ "High financial difficulty",
      iafcon %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    financial_difficulty = factor(
      financial_difficulty,
      levels = c(
        "Low financial difficulty",
        "Moderate financial difficulty",
        "High financial difficulty"
      ),
      ordered = TRUE
    )
  )
```


# 5. Physical health variables

Self-rated health, limiting long-standing illness, chronic pain, eyesight, and hearing were recoded before deriving the physical health feature set.

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    self_rated_health = case_when(
      hehelf == 1 ~ "Excellent",
      hehelf == 2 ~ "Very good",
      hehelf == 3 ~ "Good",
      hehelf == 4 ~ "Fair",
      hehelf == 5 ~ "Poor",
      hehelf %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    self_rated_health = factor(
      self_rated_health,
      levels = c(
        "Excellent",
        "Very good",
        "Good",
        "Fair",
        "Poor"
      ),
      ordered = TRUE
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    long_standing_illness = case_when(
      llsill == 0 ~ "No long-standing illness",
      llsill == 1 ~ "Long-standing illness, not limiting",
      llsill == 2 ~ "Limiting long-standing illness",
      llsill %in% c(-9, -8, -3, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    long_standing_illness = factor(
      long_standing_illness,
      levels = c(
        "No long-standing illness",
        "Long-standing illness, not limiting",
        "Limiting long-standing illness"
      ),
      ordered = TRUE
    )
  )
```


```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    chronic_pain = case_when(
      hepain == 2 ~ "No chronic pain",
      hepain == 1 & hepaa == 1 ~ "Mild pain",
      hepain == 1 & hepaa == 2 ~ "Moderate pain",
      hepain == 1 & hepaa == 3 ~ "Severe pain",
      hepain %in% c(-9, -8, -1) ~ NA_character_,
      hepain == 1 & hepaa %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    chronic_pain = factor(
      chronic_pain,
      levels = c(
        "No chronic pain",
        "Mild pain",
        "Moderate pain",
        "Severe pain"
      ),
      ordered = TRUE
    )
  )
```


```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    eyesight_problem = case_when(
      heeye %in% c(1, 2, 3) ~ "Good eyesight",
      heeye == 4 ~ "Fair eyesight",
      heeye %in% c(5, 6) ~ "Poor eyesight/blind",
      heeye %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    eyesight_problem = factor(
      eyesight_problem,
      levels = c(
        "Good eyesight",
        "Fair eyesight",
        "Poor eyesight/blind"
      ),
      ordered = TRUE
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    hearing_problem = case_when(
      hehear %in% c(1, 2, 3) ~ "Good hearing",
      hehear == 4 ~ "Fair hearing",
      hehear == 5 ~ "Poor hearing",
      hehear %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    hearing_problem = factor(
      hearing_problem,
      levels = c(
        "Good hearing",
        "Fair hearing",
        "Poor hearing"
      ),
      ordered = TRUE
    )
  )
```


# 6. Chronic disease indicators

Disease indicators combine self-reported diagnosis and follow-up disease fields where available, so each condition has one analysis-ready binary variable.
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    arthritis = case_when(
      hediadar == 1 | hedids3 == 1 | hehavear %in% c(1, 2) ~ "Yes",
      
      hediadar == 2 | hedids3 == 0 | hehavear == 3 ~ "No",
      
      hediadar %in% c(-9, -8, -1) &
        hedids3 %in% c(-9, -8, -1) &
        hehavear %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    arthritis = factor(
      arthritis,
      levels = c("No", "Yes")
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    diabetes = case_when(
      hediaddi == 1 | hedids12 == 1 | hehavedi %in% c(1, 2) ~ "Yes",
      
      hediaddi == 2 | hedids12 == 0 | hehavedi == 3 ~ "No",
      
      hediaddi %in% c(-9, -8, -1) &
        hedids12 %in% c(-9, -8, -1) &
        hehavedi %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    diabetes = factor(
      diabetes,
      levels = c("No", "Yes")
    )
  )
```

```{r}
cvd_previous_vars <- c(
  "hediacah",  # previous diagnosis: abnormal heart rhythm
  "hediacan",  # previous diagnosis: angina
  "hediachf",  # previous diagnosis: congestive heart failure
  "hediacmi",  # previous diagnosis: heart attack
  "hediacoh"   # previous diagnosis: other heart condition
)

cvd_new_vars <- c(
  "hediia95",  # new diagnosis: any other heart condition
  "hediiam2",  # new diagnosis: angina
  "hediiam3",  # new diagnosis: heart attack
  "hediiam4",  # new diagnosis: congestive heart failure
  "hediiam6"   # new diagnosis: abnormal heart rhythm
)

cvd_current_vars <- c(
  "hehaveah",  # still has abnormal heart rhythm
  "hehavean",  # still has angina
  "hehavehf",  # still has congestive heart failure
  "hehaveoh"   # still has other heart disease
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    cvd_yes_signal = 
      rowSums(across(all_of(cvd_previous_vars), ~ .x == 1), na.rm = TRUE) +
      rowSums(across(all_of(cvd_new_vars), ~ .x == 1), na.rm = TRUE) +
      rowSums(across(all_of(cvd_current_vars), ~ .x %in% c(1, 2)), na.rm = TRUE),
    
    cvd_no_signal =
      rowSums(across(all_of(cvd_previous_vars), ~ .x == 2), na.rm = TRUE) +
      rowSums(across(all_of(cvd_new_vars), ~ .x == 0), na.rm = TRUE) +
      rowSums(across(all_of(cvd_current_vars), ~ .x == 3), na.rm = TRUE),
    
    cvd = case_when(
      cvd_yes_signal > 0 ~ "Yes",
      cvd_yes_signal == 0 & cvd_no_signal > 0 ~ "No",
      TRUE ~ NA_character_
    ),
    
    cvd = factor(
      cvd,
      levels = c("No", "Yes")
    )
  ) %>%
  select(-cvd_yes_signal, -cvd_no_signal)
```


```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    stroke_yes_signal = 
      as.integer(hediacst == 1) +
      as.integer(hediiam7 == 1) +
      as.integer(hemda1 == 1) +
      as.integer(heothst == 1) +
      as.integer(hepbs == 1),
    
    stroke = case_when(
      stroke_yes_signal > 0 ~ "Yes",
      stroke_yes_signal == 0 & hediacst == 2 ~ "No",
      TRUE ~ NA_character_
    ),
    
    stroke = factor(
      stroke,
      levels = c("No", "Yes")
    )
  ) %>%
  select(-stroke_yes_signal)
```



```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    cancer = case_when(
      hediadca == 1 | hedids5 == 1 | hehaveca %in% c(1, 2) ~ "Yes",
      
      hediadca == 2 | hedids5 == 0 ~ "No",
      
      hediadca %in% c(-9, -8, -1) &
        hedids5 %in% c(-9, -8, -1) &
        hehaveca %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    cancer = factor(
      cancer,
      levels = c("No", "Yes")
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    chronic_lung_disease = case_when(
      hediadcl == 1 | hedids1 == 1 | hehavecl %in% c(1, 2) ~ "Yes",
      
      hediadcl == 2 | hedids1 == 0 | hehavecl == 3 ~ "No",
      
      hediadcl %in% c(-9, -8, -1) &
        hedids1 %in% c(-9, -8, -1) &
        hehavecl %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    chronic_lung_disease = factor(
      chronic_lung_disease,
      levels = c("No", "Yes")
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    hypertension = case_when(
      hediacbp == 1 | hediiam1 == 1 | hehavebp %in% c(1, 2) ~ "Yes",
      
      hediacbp == 2 | hediiam1 == 0 | hehavebp == 3 ~ "No",
      
      hediacbp %in% c(-9, -8, -1) &
        hediiam1 %in% c(-9, -8, -1) &
        hehavebp %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    hypertension = factor(
      hypertension,
      levels = c("No", "Yes")
    )
  )
```


```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    asthma = case_when(
      hediadas == 1 | hedids2 == 1 | hehaveas %in% c(1, 2) ~ "Yes",
      
      hediadas == 2 | hedids2 == 0 | hehaveas == 3 ~ "No",
      
      hediadas %in% c(-9, -8, -1) &
        hedids2 %in% c(-9, -8, -1) &
        hehaveas %in% c(-9, -8, -1) ~ NA_character_,
      
      TRUE ~ NA_character_
    ),
    asthma = factor(
      asthma,
      levels = c("No", "Yes")
    )
  )
```

```{r}
dementia_previous_vars <- c(
  "hediadad",  # previous diagnosis of Alzheimer's disease
  "hediaddm"   # previous diagnosis of dementia
)

dementia_new_vars <- c(
  "hedids8",   # new diagnosis: Alzheimer's disease
  "hedids9"    # new diagnosis: dementia
)

dementia_current_vars <- c(
  "hehavead",  # still has Alzheimer's disease
  "hehavedm"   # still has dementia
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    dementia_yes_signal =
      rowSums(across(all_of(dementia_previous_vars), ~ .x == 1), na.rm = TRUE) +
      rowSums(across(all_of(dementia_new_vars), ~ .x == 1), na.rm = TRUE) +
      rowSums(across(all_of(dementia_current_vars), ~ .x %in% c(1, 2)), na.rm = TRUE),
    
    dementia_no_signal =
      rowSums(across(all_of(dementia_previous_vars), ~ .x == 2), na.rm = TRUE) +
      rowSums(across(all_of(dementia_new_vars), ~ .x == 0), na.rm = TRUE) +
      rowSums(across(all_of(dementia_current_vars), ~ .x == 3), na.rm = TRUE),
    
    dementia_alzheimers = case_when(
      dementia_yes_signal > 0 ~ "Yes",
      dementia_yes_signal == 0 & dementia_no_signal > 0 ~ "No",
      TRUE ~ NA_character_
    ),
    
    dementia_alzheimers = factor(
      dementia_alzheimers,
      levels = c("No", "Yes")
    )
  ) %>%
  select(-dementia_yes_signal, -dementia_no_signal)
```



# 7. Functional limitation	
ADL count (Activities of Daily Living), IADL count(Instrumental Activities of Daily Living), mobility limitation count, falls / fear of falling if available,receiving care / receiving help
```{r}
adl_vars <- c(
  "headlba",
  "headlbe",
  "headldr",
  "headlea",
  "headlwa",
  "headlwc"
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(adl_vars),
      ~ if_else(.x %in% c(-9, -8, -3, -2, -1), NA_real_, as.numeric(.x))
    ),
    
    adl_count = rowSums(
      across(all_of(adl_vars), ~ .x == 1),
      na.rm = TRUE
    ),
    
    adl_count = if_else(
      if_all(all_of(adl_vars), is.na),
      NA_integer_,
      as.integer(adl_count)
    ),
    
    any_adl_limitation = case_when(
      is.na(adl_count) ~ NA_character_,
      adl_count == 0 ~ "No",
      adl_count >= 1 ~ "Yes"
    ),
    
    any_adl_limitation = factor(
      any_adl_limitation,
      levels = c("No", "Yes")
    )
  )
```

```{r}
iadl_vars <- c(
  "headlda",  # recognising physical danger
  "headlho",  # doing work around house and garden
  "headlma",  # using map
  "headlme",  # taking medications
  "headlmo",  # managing money
  "headlph",  # making telephone calls
  "headlpr",  # preparing hot meal
  "headlsh",  # shopping for groceries
  "headlsp"   # communication
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(iadl_vars),
      ~ if_else(.x %in% c(-9, -8, -3, -2, -1), NA_real_, as.numeric(.x))
    ),
    
    iadl_count = rowSums(
      across(all_of(iadl_vars), ~ .x == 1),
      na.rm = TRUE
    ),
    
    iadl_count = if_else(
      if_all(all_of(iadl_vars), is.na),
      NA_integer_,
      as.integer(iadl_count)
    )
  )
```


```{r}
mobility_vars <- c(
  "hemobch",  # getting up from chair
  "hemobcl",  # climbing one flight of stairs
  "hemobcs",  # climbing several flights of stairs
  "hemobli",  # lifting or carrying weights over 10 pounds
  "hemobpi",  # picking up 5p coin from table
  "hemobpu",  # pulling or pushing large objects
  "hemobre",  # reaching above shoulder level
  "hemobsi",  # sitting 2 hours
  "hemobst",  # stooping, kneeling or crouching
  "hemobwa"   # walking 100 yards
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(mobility_vars),
      ~ if_else(.x %in% c(-9, -8, -3, -2, -1), NA_real_, as.numeric(.x))
    ),
    
    mobility_count = rowSums(
      across(all_of(mobility_vars), ~ .x == 1),
      na.rm = TRUE
    ),
    
    mobility_count = if_else(
      if_all(all_of(mobility_vars), is.na),
      NA_integer_,
      as.integer(mobility_count)
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    any_iadl_limitation = case_when(
      is.na(iadl_count) ~ NA_character_,
      iadl_count == 0 ~ "No",
      iadl_count >= 1 ~ "Yes"
    ),
    any_mobility_limitation = case_when(
      is.na(mobility_count) ~ NA_character_,
      mobility_count == 0 ~ "No",
      mobility_count >= 1 ~ "Yes"
    ),
    
    any_iadl_limitation = factor(any_iadl_limitation, levels = c("No", "Yes")),
    any_mobility_limitation = factor(any_mobility_limitation, levels = c("No", "Yes"))
  )
```


```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    fall_count = case_when(
      hefla == 2 ~ 0,
      hefla == 1 & heflb >= 0 ~ as.numeric(heflb),
      hefla %in% c(-9, -8, -1) ~ NA_real_,
      hefla == 1 & heflb %in% c(-9, -8, -1) ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    fall_injury_status = case_when(
      hefla == 2 ~ "No fall",
      hefla == 1 & heflc == 1 ~ "Fall with serious injury",
      hefla == 1 & heflc == 2 ~ "Fall without serious injury",
      hefla %in% c(-9, -8, -1) ~ NA_character_,
      hefla == 1 & heflc %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    
    fall_injury_status = factor(
      fall_injury_status,
      levels = c(
        "No fall",
        "Fall without serious injury",
        "Fall with serious injury"
      ),
      ordered = TRUE
    )
  )
```



## 8. Mental health	
depressive symptoms / CES-D score (Center for Epidemiologic Studies Depression Scale), sleep problems, anxiety symptoms / GAD-7 score
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    cesd_score = case_when(
      cesd_na == 8 & cesd_sc >= 0 ~ as.numeric(cesd_sc),
      cesd_na < 8 ~ NA_real_,
      cesd_sc %in% c(-1, -2) ~ NA_real_,
      TRUE ~ NA_real_
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    restless_sleep = case_when(
      pscedc == 1 ~ "Yes",
      pscedc == 2 ~ "No",
      pscedc %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    restless_sleep = factor(
      restless_sleep,
      levels = c("No", "Yes")
    )
  )
```


```{r}
gad7_vars <- c(
  "scafrai",  # afraid something awful might happen
  "scanno",   # easily annoyed or irritable
  "scanxe",   # anxious or on edge
  "scresl",   # restless
  "scrlax",   # trouble relaxing
  "scwrys",   # not able to stop/control worrying
  "scwryt"    # worried too much about different things
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(gad7_vars),
      ~ case_when(
        .x %in% c(-9, -8, -2, -1) ~ NA_real_,
        .x %in% c(1, 2, 3, 4) ~ as.numeric(.x) - 1,
        TRUE ~ NA_real_
      ),
      .names = "{.col}_score"
    )
  )
```

```{r}
gad7_score_vars <- paste0(gad7_vars, "_score")

data11_selected_features <- data11_selected_features %>%
  mutate(
    gad7_valid_n = rowSums(
      !is.na(across(all_of(gad7_score_vars)))
    ),
    
    anxiety_score = case_when(
      gad7_valid_n == 7 ~ rowSums(across(all_of(gad7_score_vars))),
      gad7_valid_n < 7 ~ NA_real_
    )
  )
```



## 9. Loneliness	
loneliness score
```{r}
loneliness_vars <- c("scfeela", "scfeelb", "scfeelc")

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(loneliness_vars),
      ~ if_else(.x %in% c(-9, -8, -2, -1), NA_real_, as.numeric(.x))
    ),
    
    loneliness_score = if_else(
      if_any(all_of(loneliness_vars), is.na),
      NA_real_,
      rowSums(across(all_of(loneliness_vars)))
    )
  )
```


## 10. Social participation	
social participation count, volunteering participation, internet use, social support / social relationship quality, caring responsibilities, food insecurity, subjective financial strain/subjective social status
```{r}
social_org_vars <- c(
  "scorg95",  # other organisations/clubs/societies
  "scorgch",  # charitable associations
  "scorged",  # education, arts, music groups/evening classes
  "scorgnw",  # tenants/resident/neighbourhood watch groups
  "scorgpo",  # political party/trade union/environmental groups
  "scorgrl",  # church or religious group
  "scorgsc",  # social clubs
  "scorgsp"   # sports clubs/gyms/exercise classes
)

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(c(social_org_vars, "scorg96")),
      ~ if_else(.x %in% c(-9, -8, -2, -1), NA_real_, as.numeric(.x))
    ),
    
    social_participation_count = rowSums(
      across(all_of(social_org_vars), ~ .x == 1),
      na.rm = TRUE
    ),
    
    social_participation_count = case_when(
      scorg96 == 1 ~ 0L,
      
      if_all(all_of(social_org_vars), is.na) ~ NA_integer_,
      
      TRUE ~ as.integer(social_participation_count)
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    volunteering_participation = case_when(
      erfvoft == 1 ~ "At least once a week",
      erfvoft == 2 ~ "At least once a month",
      erfvoft == 3 ~ "Less often",
      erfvoft == 4 ~ "One-off activity",
      erfvoft %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    volunteering_participation = factor(
      volunteering_participation,
      levels = c(
        "One-off activity",
        "Less often",
        "At least once a month",
        "At least once a week"
      ),
      ordered = TRUE
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    internet_help_frequency = case_when(
      scinf == 1 ~ "Help all of the time",
      scinf == 2 ~ "Help some of the time",
      scinf == 3 ~ "Rarely or never help",
      scinf %in% c(-9, -8, -2, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    internet_help_frequency = factor(
      internet_help_frequency,
      levels = c(
        "Rarely or never help",
        "Help some of the time",
        "Help all of the time"
      ),
      ordered = TRUE
    )
  )
```

```{r}
# social relationship quality variables

child_positive_vars <- c("scchda", "scchdb", "scchdc")
child_negative_vars <- c("scchdd")

family_positive_vars <- c("scfama", "scfamb", "scfamc")
family_negative_vars <- c("scfamd")

friend_positive_vars <- c("scfrda", "scfrdb", "scfrdc")
friend_negative_vars <- c("scfrdd")

partner_positive_vars <- c("scprta", "scprtb", "scprtc")
partner_negative_vars <- c("scprtd")

positive_support_vars <- c(
  child_positive_vars,
  family_positive_vars,
  friend_positive_vars,
  partner_positive_vars
)

negative_relationship_vars <- c(
  child_negative_vars,
  family_negative_vars,
  friend_negative_vars,
  partner_negative_vars
)

social_relationship_vars <- c(
  positive_support_vars,
  negative_relationship_vars
)
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(social_relationship_vars),
      ~ if_else(.x %in% c(-9, -8, -2, -1), NA_real_, as.numeric(.x))
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(positive_support_vars),
      ~ case_when(
        .x == 1 ~ 4,
        .x == 2 ~ 3,
        .x == 3 ~ 2,
        .x == 4 ~ 1,
        TRUE ~ NA_real_
      ),
      .names = "{.col}_score"
    ),
    
    across(
      all_of(negative_relationship_vars),
      ~ case_when(
        .x == 1 ~ 1,
        .x == 2 ~ 2,
        .x == 3 ~ 3,
        .x == 4 ~ 4,
        TRUE ~ NA_real_
      ),
      .names = "{.col}_score"
    )
  )
```

```{r}
social_relationship_score_vars <- paste0(social_relationship_vars, "_score")

data11_selected_features <- data11_selected_features %>%
  mutate(
    social_relationship_quality_score = if_else(
      if_all(all_of(social_relationship_score_vars), is.na),
      NA_real_,
      rowMeans(across(all_of(social_relationship_score_vars)), na.rm = TRUE)
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    caring_responsibilities = case_when(
      ercaa == 1 ~ "Yes",
      ercaa == 2 ~ "No",
      ercaa %in% c(-9, -8, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    caring_responsibilities = factor(
      caring_responsibilities,
      levels = c("No", "Yes")
    )
  )
```


```{r}
food_insecurity_vars <- c("homeal", "scaffda", "scaffdb")

data11_selected_features <- data11_selected_features %>%
  mutate(
    food_insecurity_yes_signal =
      as.integer(homeal == 1) +
      as.integer(scaffda %in% c(1, 2)) +
      as.integer(scaffdb %in% c(1, 2)),
    
    food_insecurity_valid_n =
      as.integer(homeal %in% c(1, 2)) +
      as.integer(scaffda %in% c(1, 2, 3)) +
      as.integer(scaffdb %in% c(1, 2, 3)),
    
    food_insecurity = case_when(
      food_insecurity_yes_signal > 0 ~ "Yes",
      food_insecurity_yes_signal == 0 & food_insecurity_valid_n > 0 ~ "No",
      food_insecurity_valid_n == 0 ~ NA_character_,
      TRUE ~ NA_character_
    ),
    
    food_insecurity = factor(
      food_insecurity,
      levels = c("No", "Yes")
    )
  ) %>%
  select(
    -food_insecurity_yes_signal,
    -food_insecurity_valid_n
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    subjective_social_status = case_when(
      sclddr %in% c(-9, -8, -2, -1) ~ NA_real_,
      TRUE ~ as.numeric(sclddr)
    )
  )
```


## 11. Health behaviour	
smoking status, alcohol consumption, BMI (Body Mass Index), physical activity
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    smoking_status = case_when(
      smokerstat == 0 ~ "Never smoker",
      smokerstat %in% c(1, 2, 3) ~ "Former smoker",
      smokerstat == 4 ~ "Current smoker",
      smokerstat %in% c(-9, -8, -2) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    smoking_status = factor(
      smoking_status,
      levels = c(
        "Never smoker",
        "Former smoker",
        "Current smoker"
      )
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    alcohol_status = case_when(
      scako %in% c(1, 2, 3, 4, 5, 6, 7, 9) ~ "Drinker",
      scako == 8 ~ "Non-drinker",
      scako %in% c(-9, -8, -2, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    alcohol_status = factor(
      alcohol_status,
      levels = c("Non-drinker", "Drinker")
    ),
    
    alcohol_frequency_group = case_when(
      scako == 8 ~ "No alcohol in last 12 months",
      scako == 7 ~ "Rarely",
      scako %in% c(5, 6) ~ "Monthly or less",
      scako %in% c(2, 3, 4) ~ "Weekly",
      scako == 1 ~ "Almost every day",
      scako == 9 ~ NA_character_,
      scako %in% c(-9, -8, -2, -1) ~ NA_character_,
      TRUE ~ NA_character_
    ),
    alcohol_frequency_group = factor(
      alcohol_frequency_group,
      levels = c(
        "No alcohol in last 12 months",
        "Rarely",
        "Monthly or less",
        "Weekly",
        "Almost every day"
      ),
      ordered = TRUE
    )
  )
```

```{r}
physical_activity_vars <- c("heacta", "heactb", "heactc")

data11_selected_features <- data11_selected_features %>%
  mutate(
    across(
      all_of(physical_activity_vars),
      ~ if_else(.x %in% c(-9, -8, -1), NA_real_, as.numeric(.x))
    ),
    
    physical_activity_level = case_when(
      heacta %in% c(1, 2) ~ "Vigorous weekly",
      heactb %in% c(1, 2) ~ "Moderate weekly",
      heactc %in% c(1, 2) ~ "Mild weekly only",
      if_any(all_of(physical_activity_vars), ~ .x == 3) ~ "Less than weekly",
      if_all(all_of(physical_activity_vars), ~ .x == 4) ~ "Inactive",
      TRUE ~ NA_character_
    ),
    
    physical_activity_level = factor(
      physical_activity_level,
      levels = c(
        "Inactive",
        "Less than weekly",
        "Mild weekly only",
        "Moderate weekly",
        "Vigorous weekly"
      ),
      ordered = TRUE
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    bmival_clean = if_else(
      bmival %in% c(-9, -8, -1),
      NA_real_,
      as.numeric(bmival)
    ),
    
    bmi_clean = if_else(
      bmi %in% c(-9, -8, -1),
      NA_real_,
      as.numeric(bmi)
    ),
    
    bmi_new = coalesce(bmival_clean, bmi_clean)
  )
```


## 12. Cognition	
memory score, orientation score
```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    memory_score = case_when(
      memtotb %in% c(-9, -3, -2, -1) ~ NA_real_,
      TRUE ~ as.numeric(memtotb)
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    orientation_score = case_when(
      cfdscr %in% c(-9, -8, -1) ~ NA_real_,
      TRUE ~ as.numeric(cfdscr)
    )
  )
```

```{r}
data11_selected_features <- data11_selected_features %>%
  mutate(
    receiving_help_count = case_when(
      catno %in% c(-9, -8, -1) ~ NA_real_,
      TRUE ~ as.numeric(catno)
    )
  )
```


Chronic disease count is not constructed at this stage. The individual disease indicators are kept for the pre-screening missingness check first; the disease count can be rebuilt later after high-missingness disease variables have been excluded.





# 13. Save base modelling data

At this stage, only the base features from the preliminary feature-selection table are retained. 

```{r}
base_feature_vars <- c(
  "age",
  "sex",
  "education_level",
  "marital_status",
  "wealth_quintile",
  "dhwork",
  "housing_tenure",
  "financial_difficulty",
  "self_rated_health",
  "long_standing_illness",
  "chronic_pain",
  "eyesight_problem",
  "hearing_problem",
  "arthritis",
  "diabetes",
  "cvd",
  "stroke",
  "cancer",
  "chronic_lung_disease",
  "hypertension",
  "dementia_alzheimers",
  "asthma",
  "adl_count",
  "iadl_count",
  "mobility_count",
  "fall_injury_status",
  "cesd_score",
  "restless_sleep",
  "loneliness_score",
  "social_participation_count",
  "volunteering_participation",
  "internet_help_frequency",
  "social_relationship_quality_score",
  "caring_responsibilities",
  "smoking_status",
  "alcohol_frequency_group",
  "physical_activity_level",
  "bmi_new",
  "memory_score",
  "orientation_score",
  "food_insecurity",
  "subjective_social_status",
  "anxiety_score",
  "receiving_help_count"
)

base_feature_vars
```

```{r}
data11_base_features <- data11_selected_features %>%
  select(
    idauniq,
    idahhw11,
    perid,
    casp19_total,
    all_of(base_feature_vars)
  ) %>%
  mutate(
    qol_cutoff = 35,
    qol_status = case_when(
      casp19_total <= qol_cutoff ~ "Low QoL",
      casp19_total > qol_cutoff ~ "Non-low QoL",
      TRUE ~ NA_character_
    ),
    qol_status = factor(
      qol_status,
      levels = c("Non-low QoL", "Low QoL")
    )
  ) %>%
  select(
    idauniq,
    idahhw11,
    perid,
    casp19_total,
    qol_status,
    all_of(base_feature_vars)
  )

dim(data11_base_features)
```

```{r}
data11_base_features
```
The base modelling dataset contains 5,314 respondents and 49 columns. These include three identifier columns (`idauniq`, `idahhw11`, and `perid`), two outcome-related columns (`casp19_total` and `qol_status`), and 44 base candidate predictor features. At this stage, the dataset keeps only the initial base feature set; additional candidate features will be searched and added in the next step.

```{r}
saveRDS(
  data11_base_features,
  file.path(processed_path, "07_wave11_ml_base_features.rds")
)

saveRDS(
  base_feature_vars,
  file.path(processed_path, "07_base_candidate_features.rds")
)

message("Saved: data/processed/07_wave11_ml_base_features.rds")
message("Saved: data/processed/07_base_candidate_features.rds")
```
