---
title: "07_EDA02_additional_features"
author: "Yiwen Xiang"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: true
---

# 1. Setup

```{r, include=FALSE}
source(here::here("script", "00_packages.R"))
source(here::here("script", "02_functions.R"))

w11_datasets <- readRDS(file.path(processed_path, "01_wave11_datasets.rds"))
data11_main <- w11_datasets$w11_main
data11_ifs <- w11_datasets$w11_ifs
data11_financial <- w11_datasets$w11_financial
data11_nurse <- w11_datasets$w11_nurse
var_w11_all_labeled <- readRDS(file.path(processed_path, "03var_w11_all_labeled.rds"))
```

```{r}
data11_analysis_outcome_valid <- readRDS(
  file.path(processed_path, "06_wave11_outcome.rds")
)

head(data11_analysis_outcome_valid)
```

```{r}
dim(data11_analysis_outcome_valid)
```

```{r}
# Search across all available Wave 11 tables
candidate_tables <- unique(var_w11_all_labeled$table)

candidate_tables
```

# 2. Check candidate missingness after searching
```{r}
check_candidate_missingness <- function(search_result, analysis_sample) {
  search_result %>%
    dplyr::filter(table %in% names(w11_datasets)) %>%
    dplyr::distinct(table, variable, .keep_all = TRUE) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      n_analysis = nrow(analysis_sample),
      n_missing_analysis = {
        df <- w11_datasets[[table]]
        
        if (!variable %in% names(df)) {
          NA_integer_
        } else {
          temp <- analysis_sample %>%
            dplyr::select(idauniq) %>%
            dplyr::left_join(
              df %>%
                dplyr::select(
                  idauniq,
                  value = dplyr::all_of(variable)
                ),
              by = "idauniq"
            )
          
          x <- temp$value
          
          # ELSA negative values are treated as special missing codes.
          if (is.numeric(x) | is.integer(x)) {
            sum(is.na(x) | x < 0)
          } else {
            sum(is.na(x))
          }
        }
      },
      missing_pct_analysis = round(
        n_missing_analysis / n_analysis * 100,
        2
      )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(missing_pct_analysis, table, variable)
}
```

# 3. Subjective social status
```{r}
subjective_social_status_search <- search_vars(
  keywords = c(
    "subjective social status",
    "social status",
    "social ladder",
    "ladder",
    "social standing",
    "social position",
    "where would you place yourself",
    "best off",
    "worst off",
    "sclddr"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

subjective_social_status_search_checked <- check_candidate_missingness(
  subjective_social_status_search,
  data11_analysis_outcome_valid
)

subjective_social_status_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(30)
```


# 4. Any psychiatric problem
Psychiatric problem fields are searched before constructing `any_psychiatric_problem`.
```{r}
psychiatric_problem_search <- search_vars(
  keywords = c(
    "psychiatric problem",
    "psychiatric",
    "mental problem",
    "mental health problem",
    "emotional problem",
    "emotional problems",
    "hallucinations",
    "hallucination",
    "anxiety",
    "depression",
    "depressed",
    "schizophrenia",
    "psychosis",
    "psychotic",
    "mood swings",
    "manic depression",
    "bipolar",
    "nervous breakdown",
    "other psychiatric"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

psychiatric_problem_search_checked <- check_candidate_missingness(
  psychiatric_problem_search,
  data11_analysis_outcome_valid
)

psychiatric_problem_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(80)
```

# 5. Hypertension
```{r}
hypertension_search <- search_vars(
  keywords = c(
    "hypertension",
    "high blood pressure",
    "blood pressure",
    "raised blood pressure",
    "doctor diagnosed high blood pressure",
    "ever had high blood pressure",
    "has high blood pressure",
    "bp",
    "hibpe",
    "hehibp",
    "hibp",
    "hypert"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

hypertension_search_checked <- check_candidate_missingness(
  hypertension_search,
  data11_analysis_outcome_valid
)

hypertension_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(50)
```

# 6. Falls / fear of falling
```{r}
falls_fear_search <- search_vars(
  keywords = c(
    "fall",
    "falls",
    "fallen",
    "falling",
    "fear of falling",
    "afraid of falling",
    "worried about falling",
    "worry about falling",
    "concerned about falling",
    "unsteady",
    "balance",
    "trip",
    "slip",
    "fracture",
    "hip fracture"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

falls_fear_search_checked <- check_candidate_missingness(
  falls_fear_search,
  data11_analysis_outcome_valid
)

falls_fear_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(80)
```

# 7. Volunteering participation
```{r}
volunteering_search <- search_vars(
  keywords = c(
    "volunteer",
    "volunteering",
    "voluntary",
    "voluntary work",
    "voluntary organisation",
    "voluntary organization",
    "charity",
    "charitable",
    "helping others",
    "unpaid work",
    "community work",
    "community group",
    "civic participation",
    "social participation"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

volunteering_search_checked <- check_candidate_missingness(
  volunteering_search,
  data11_analysis_outcome_valid
)

volunteering_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(80)
```

# 8. Memory score / cognition
```{r}
memory_score_search <- search_vars(
  keywords = c(
    "memory",
    "memory score",
    "immediate recall",
    "delayed recall",
    "word recall",
    "recall",
    "orientation",
    "orientation score",
    "cognitive",
    "cognition",
    "cognitive function",
    "mental status",
    "date naming",
    "day",
    "month",
    "year",
    "serial",
    "numeracy",
    "animal naming",
    "verbal fluency"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

memory_score_search
```

```{r}
id_variables <- c(
  "idauniq",
  "idahhw11",
  "idahh",
  "perid",
  "hhid",
  "pid",
  "serial",
  "wave"
)

memory_score_search_clean <- memory_score_search %>%
  filter(!variable %in% id_variables)

memory_score_search_checked <- check_candidate_missingness(
  memory_score_search_clean,
  data11_analysis_outcome_valid
)

memory_score_search_checked %>%
  select(
    table,
    variable,
    label,
    type,
    n_missing_analysis,
    missing_pct_analysis
  ) %>%
  head(100)
```

# 9. Receiving care / receiving help
```{r}
receiving_care_help_search <- search_vars(
  keywords = c(
    "receive care",
    "receiving care",
    "received care",
    "get care",
    "getting care",
    "got care",
    "receive help",
    "receiving help",
    "received help",
    "get help",
    "getting help",
    "got help",
    "help received",
    "help with activities",
    "help with daily activities",
    "help with personal care",
    "personal care",
    "home care",
    "social care",
    "care package",
    "care services",
    "social services",
    "formal care",
    "informal care",
    "paid care",
    "paid carer",
    "paid helper",
    "carer",
    "caregiver",
    "helper",
    "home help",
    "meals on wheels",
    "nursing care",
    "residential care",
    "sheltered housing",
    "support worker",
    "received assistance",
    "assistance",
    "help from"
  ),
  var_dict = var_w11_all_labeled
) %>%
  filter(table %in% candidate_tables)

receiving_care_help_search_checked <- check_candidate_missingness(
  receiving_care_help_search,
  data11_analysis_outcome_valid
)

receiving_care_help_search_checked %>%
  select(table, variable, label, type, n_missing_analysis, missing_pct_analysis) %>%
  head(100)
```

# 10. Combine search results
```{r}
additional_candidate_search_results <- bind_rows(
  subjective_social_status_search_checked %>%
    mutate(search_topic = "subjective_social_status"),
  
  psychiatric_problem_search_checked %>%
    mutate(search_topic = "psychiatric_problem"),
  
  hypertension_search_checked %>%
    mutate(search_topic = "hypertension"),
  
  falls_fear_search_checked %>%
    mutate(search_topic = "falls_fear_of_falling"),
  
  volunteering_search_checked %>%
    mutate(search_topic = "volunteering"),
  
  memory_score_search_checked %>%
    mutate(search_topic = "memory_cognition"),
  
  receiving_care_help_search_checked %>%
    mutate(search_topic = "receiving_care_help")
) %>%
  select(
    search_topic,
    table,
    variable,
    label,
    type,
    n_analysis,
    n_missing_analysis,
    missing_pct_analysis,
    everything()
  ) %>%
  distinct(search_topic, table, variable, .keep_all = TRUE) %>%
  arrange(search_topic, missing_pct_analysis, table, variable)

additional_candidate_search_results
```

Candidate fields with missingness not exceeding 20% were retained for manual review:
```{r}
additional_candidate_under20 <- additional_candidate_search_results %>%
  filter(
    !is.na(missing_pct_analysis),
    missing_pct_analysis <= 20
  ) %>%
  arrange(search_topic, missing_pct_analysis)

additional_candidate_under20
```

# 11. Export search results
```{r}
export_xlsx(
  additional_candidate_under20,
  "additional_candidate_under20.xlsx",
  output_path
)
```

# 12. Construct additional candidate features

For concepts that overlap with base features, the table below documents whether the additional candidate uses alternative or expanded source columns.

```{r}
additional_feature_source_review <- tibble::tribble(
  ~concept, ~base_feature, ~base_source_columns, ~additional_feature, ~additional_source_columns, ~note,
  "Falls", "fall_injury_status", "hefla, heflc", "fall_status", "hefla, heflc, hefrac", "The additional version uses an expanded fall-related source set, including hip fracture information.",
  "Hypertension", "hypertension", "hediacbp, hediiam1, hehavebp", "hypertension_status", "hediacbp, hediiam1, hediia1, hehavebp", "The additional version uses an expanded diagnosis source set and a more explicit hypertension status definition.",
  "Memory", "memory_score", "memtotb", "objective_memory_score; self_rated_memory; memory_change", "cflisen, cflisd, cfmetm, cfmetmt", "The additional memory features use different cognitive and self-rated memory fields.",
  "Care/help", "receiving_help_count", "catno", "any_received_care_help", "cahsmpc and cahn*/cahs*/caany care-help indicators", "The additional version uses detailed care/help source indicators instead of only the count field."
)

additional_feature_source_review
```

An ID-only table was created first so that each newly constructed additional feature could be joined back by respondent ID.
```{r}
data11_additional_features <- data11_analysis_outcome_valid %>%
  select(idauniq)

data11_additional_features
```

```{r}
fall_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    hefla,
    heflc,
    hefrac
  ) %>%
  mutate(
    hefla_clean = case_when(
      is.na(hefla) ~ NA_real_,
      as.numeric(hefla) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hefla)
    ),
    
    heflc_clean = case_when(
      is.na(heflc) ~ NA_real_,
      as.numeric(heflc) < 0 ~ NA_real_,
      TRUE ~ as.numeric(heflc)
    ),
    
    hefrac_clean = case_when(
      is.na(hefrac) ~ NA_real_,
      as.numeric(hefrac) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hefrac)
    ),
    
    fall_status = case_when(
      hefrac_clean == 1 | heflc_clean == 1 ~ "Fall with serious injury or hip fracture",
      hefla_clean == 1 & heflc_clean == 2 ~ "Fall without serious injury",
      hefla_clean == 2 ~ "No recent fall",
      TRUE ~ NA_character_
    ),
    
    fall_status = factor(
      fall_status,
      levels = c(
        "No recent fall",
        "Fall without serious injury",
        "Fall with serious injury or hip fracture"
      ),
      ordered = TRUE
    )
  ) %>%
  select(
    idauniq,
    fall_status
  )

data11_additional_features <- data11_additional_features %>%
  select(-any_of("fall_status")) %>%
  left_join(fall_feature, by = "idauniq")

```

```{r}
hypertension_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    hediacbp,
    hediiam1,
    hediia1,
    hehavebp
  ) %>%
  mutate(
    hediacbp_clean = case_when(
      is.na(hediacbp) ~ NA_real_,
      as.numeric(hediacbp) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hediacbp)
    ),
    
    hediiam1_clean = case_when(
      is.na(hediiam1) ~ NA_real_,
      as.numeric(hediiam1) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hediiam1)
    ),
    
    hediia1_clean = case_when(
      is.na(hediia1) ~ NA_real_,
      as.numeric(hediia1) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hediia1)
    ),
    
    hehavebp_clean = case_when(
      is.na(hehavebp) ~ NA_real_,
      as.numeric(hehavebp) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hehavebp)
    ),
    
    new_hypertension_diagnosis = coalesce(
      hediiam1_clean,
      hediia1_clean
    ),
    
    hypertension_status = case_when(
      hediacbp_clean == 1 |
        new_hypertension_diagnosis == 1 |
        hehavebp_clean == 1 ~ "Hypertension",
      
      hediacbp_clean == 2 &
        new_hypertension_diagnosis == 2 &
        (is.na(hehavebp_clean) | hehavebp_clean == 2) ~ "No hypertension",
      
      TRUE ~ NA_character_
    ),
    
    hypertension_status = factor(
      hypertension_status,
      levels = c(
        "No hypertension",
        "Hypertension"
      )
    )
  ) %>%
  select(
    idauniq,
    hypertension_status
  )

data11_additional_features <- data11_additional_features %>%
  select(-any_of("hypertension_status")) %>%
  left_join(hypertension_feature, by = "idauniq")

```



```{r}
memory_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    cflisen,
    cflisd,
    cfmetm,
    cfmetmt
  ) %>%
  mutate(
    cflisen_clean = case_when(
      is.na(cflisen) ~ NA_real_,
      as.numeric(cflisen) < 0 ~ NA_real_,
      TRUE ~ as.numeric(cflisen)
    ),
    
    cflisd_clean = case_when(
      is.na(cflisd) ~ NA_real_,
      as.numeric(cflisd) < 0 ~ NA_real_,
      TRUE ~ as.numeric(cflisd)
    ),
    
    cfmetm_clean = case_when(
      is.na(cfmetm) ~ NA_real_,
      as.numeric(cfmetm) < 0 ~ NA_real_,
      TRUE ~ as.numeric(cfmetm)
    ),
    
    cfmetmt_clean = case_when(
      is.na(cfmetmt) ~ NA_real_,
      as.numeric(cfmetmt) < 0 ~ NA_real_,
      TRUE ~ as.numeric(cfmetmt)
    ),
    
    # Higher values indicate better objective memory performance.
    objective_memory_score = case_when(
      !is.na(cflisen_clean) & !is.na(cflisd_clean) ~ 
        cflisen_clean + cflisd_clean,
      TRUE ~ NA_real_
    ),
    
    self_rated_memory = case_when(
      cfmetm_clean == 1 ~ 5,
      cfmetm_clean == 2 ~ 4,
      cfmetm_clean == 3 ~ 3,
      cfmetm_clean == 4 ~ 2,
      cfmetm_clean == 5 ~ 1,
      TRUE ~ NA_real_
    ),
    
    memory_change = case_when(
      cfmetmt_clean == 1 ~ 3,
      cfmetmt_clean == 2 ~ 2,
      cfmetmt_clean == 3 ~ 1,
      TRUE ~ NA_real_
    )
  ) %>%
  select(
    idauniq,
    objective_memory_score,
    self_rated_memory,
    memory_change
  )

data11_additional_features <- data11_additional_features %>%
  select(
    -any_of(c(
      "objective_memory_score",
      "self_rated_memory",
      "memory_change"
    ))
  ) %>%
  left_join(memory_feature, by = "idauniq")

```


```{r}
psychiatric_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    hedids7
  ) %>%
  mutate(
    hedids7_clean = case_when(
      is.na(hedids7) ~ NA_real_,
      as.numeric(hedids7) < 0 ~ NA_real_,
      TRUE ~ as.numeric(hedids7)
    ),
    
    any_psychiatric_problem = case_when(
      hedids7_clean == 1 ~ "Psychiatric problem mentioned",
      hedids7_clean == 0 ~ "No psychiatric problem mentioned",
      TRUE ~ NA_character_
    ),
    
    any_psychiatric_problem = factor(
      any_psychiatric_problem,
      levels = c(
        "No psychiatric problem mentioned",
        "Psychiatric problem mentioned"
      )
    )
  ) %>%
  select(
    idauniq,
    any_psychiatric_problem
  )

data11_additional_features <- data11_additional_features %>%
  select(-any_of("any_psychiatric_problem")) %>%
  left_join(psychiatric_feature, by = "idauniq")

```


```{r}
# 0/1 variables: 1 = mentioned.
mentioned_care_vars <- c(
  "cahsmpc"
)

# 1/2 variables: 1 = yes, 2 = no.
yesno_care_vars <- c(
  # Informal help sources.
  "cahnbr",
  "cahnda",
  "cahnfr",
  "cahngc",
  "cahnhw",
  "cahnne",
  "cahnor",
  "cahnsi",
  "cahnso",
  
  # Formal help sources.
  "cahncl",
  "cahnhc",
  "cahnhm",
  "cahnnh",
  "cahnot",
  "cahnre",
  "cahnvo",
  "cahnwa",
  
  # Care home and non-family care.
  "cahsa",
  "caany"
)

care_help_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    any_of(c(mentioned_care_vars, yesno_care_vars))
  ) %>%
  mutate(
    across(
      any_of(c(mentioned_care_vars, yesno_care_vars)),
      ~ case_when(
        is.na(.) ~ NA_real_,
        as.numeric(.) < 0 ~ NA_real_,
        TRUE ~ as.numeric(.)
      ),
      .names = "{.col}_clean"
    )
  ) %>%
  rowwise() %>%
  mutate(
    any_received_care_help = {
      
      mentioned_x <- c_across(
        any_of(paste0(mentioned_care_vars, "_clean"))
      )
      
      yesno_x <- c_across(
        any_of(paste0(yesno_care_vars, "_clean"))
      )
      
      all_x <- c(mentioned_x, yesno_x)
      
      case_when(
        any(mentioned_x == 1, na.rm = TRUE) |
          any(yesno_x == 1, na.rm = TRUE) ~ "Received care/help",
        
        all(mentioned_x %in% c(0, NA), na.rm = FALSE) &
          all(yesno_x %in% c(2, NA), na.rm = FALSE) &
          any(!is.na(all_x)) ~ "No care/help received",
        
        TRUE ~ NA_character_
      )
    },
    
    any_received_care_help = factor(
      any_received_care_help,
      levels = c(
        "No care/help received",
        "Received care/help"
      )
    )
  ) %>%
  ungroup() %>%
  select(
    idauniq,
    any_received_care_help
  )

data11_additional_features <- data11_additional_features %>%
  select(-any_of("any_received_care_help")) %>%
  left_join(care_help_feature, by = "idauniq")

```

```{r}
unmet_personal_care_feature <- data11_analysis_outcome_valid %>%
  select(
    idauniq,
    capsac10
  ) %>%
  mutate(
    capsac10_clean = case_when(
      is.na(capsac10) ~ NA_real_,
      as.numeric(capsac10) < 0 ~ NA_real_,
      TRUE ~ as.numeric(capsac10)
    ),
    
    unmet_personal_care_need = case_when(
      capsac10_clean == 1 ~ "Unmet personal care need",
      capsac10_clean == 0 ~ "No unmet personal care need",
      TRUE ~ NA_character_
    ),
    
    unmet_personal_care_need = factor(
      unmet_personal_care_need,
      levels = c(
        "No unmet personal care need",
        "Unmet personal care need"
      )
    )
  ) %>%
  select(
    idauniq,
    unmet_personal_care_need
  )

data11_additional_features <- data11_additional_features %>%
  select(-any_of("unmet_personal_care_need")) %>%
  left_join(unmet_personal_care_feature, by = "idauniq")

```

```{r}
data11_analysis_outcome_valid %>%
  count(sclddr, sort = TRUE) %>%
  mutate(
    pct = round(n / sum(n) * 100, 2)
  )
```


# 13. Check additional feature missingness

Additional features were checked for missingness at this stage, but features were not excluded here. Missingness-based exclusion is applied to the complete candidate feature set in the next script.

```{r}
check_all_missing <- function(data, id_cols = c("idauniq")) {
  feature_cols <- setdiff(names(data), id_cols)
  n_total <- nrow(data)

  data %>%
    select(all_of(feature_cols)) %>%
    summarise(across(everything(), ~ sum(is.na(.x)))) %>%
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "n_missing"
    ) %>%
    mutate(
      n_total = n_total,
      n_valid = n_total - n_missing,
      missing_pct = round(n_missing / n_total * 100, 2)
    ) %>%
    select(
      variable,
      n_total,
      n_missing,
      n_valid,
      missing_pct
    ) %>%
    arrange(desc(missing_pct), variable)
}

additional_missingness <- check_all_missing(data11_additional_features)

additional_missingness
```

# 14. Merge base and additional candidate features

The additional candidate variables were searched and constructed first, then merged with the full base candidate feature table. Variables already present in the base table, such as `volunteering_participation`, are not added a second time.

```{r}
data11_base_features <- readRDS(
  file.path(processed_path, "07_wave11_ml_base_features.rds")
)

base_feature_vars <- readRDS(
  file.path(processed_path, "07_base_candidate_features.rds")
)

dim(data11_base_features)
length(base_feature_vars)
```

```{r}
additional_feature_vars <- setdiff(
  names(data11_additional_features),
  c(
    "idauniq",
    "idahhw11",
    "perid",
    names(data11_base_features)
  )
)

additional_feature_vars
```

```{r}
data11_ml_candidate_features <- data11_base_features %>%
  left_join(
    data11_additional_features %>%
      select(idauniq, all_of(additional_feature_vars)),
    by = "idauniq",
    relationship = "one-to-one"
  )

candidate_features_pre_screening <- c(
  base_feature_vars,
  additional_feature_vars
)

candidate_features_pre_screening <- intersect(
  candidate_features_pre_screening,
  names(data11_ml_candidate_features)
)
```

```{r}
data11_ml_candidate_features
```


```{r}
dim(data11_ml_candidate_features)
length(candidate_features_pre_screening)
```

The merged candidate dataset contains the three identifier columns, two outcome-related columns, and all base plus additional candidate features. These candidate features are screened together in the next script.

```{r}
saveRDS(
  data11_ml_candidate_features,
  file.path(processed_path, "07_wave11_ml_candidate_features_pre_screening.rds")
)

saveRDS(
  candidate_features_pre_screening,
  file.path(processed_path, "07_candidate_features_pre_screening.rds")
)


message("Saved: data/processed/07_wave11_ml_candidate_features_pre_screening.rds")
message("Saved: data/processed/07_candidate_features_pre_screening.rds")
```
