---
title: "09_heldout_test_evaluation"
author: "Yiwen Xiang"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# 09 Held-out test evaluation

This script evaluates the fixed models developed in `08_model_development_training_only.Rmd` on the held-out test set. No hyperparameters, predictors, class weights or classification thresholds are selected in this script.

```{r}
source(here::here("script", "00_packages.R"))
```

```{r}
data11_training <- readRDS(
  file.path(processed_path, "07_wave11_ml_data.rds")
)

model_bundle <- readRDS(
  file.path(processed_path, "08_final_model_bundle.rds")
)

split_ids <- readRDS(
  file.path(processed_path, "08_train_test_split_ids.rds")
)

id_vars <- c("idauniq", "idahhw11", "perid")
full_feature_names <- model_bundle$full_feature_names
selected_main_features <- model_bundle$selected_main_features

stopifnot(
  length(full_feature_names) == 42L,
  !anyDuplicated(full_feature_names),
  all(full_feature_names %in% names(data11_training)),
  all(selected_main_features %in% names(data11_training)),
  identical(model_bundle$split_ids$train_idauniq, split_ids$train_idauniq),
  identical(model_bundle$split_ids$test_idauniq, split_ids$test_idauniq)
)
```

```{r}
# Reconstruct the same train-test split generated in script 08.
split_train_ids <- split_ids$train_idauniq
split_test_ids <- split_ids$test_idauniq

stopifnot(!anyDuplicated(data11_training$idauniq))

train_position <- match(
  split_train_ids,
  data11_training$idauniq
)

test_position <- match(
  split_test_ids,
  data11_training$idauniq
)

if (anyNA(train_position) || anyNA(test_position)) {
  stop("Some split IDs were not found in the Wave 11 modelling dataset.")
}

data11_train <- data11_training[
  train_position,
  ,
  drop = FALSE
]

data11_test <- data11_training[
  test_position,
  ,
  drop = FALSE
]

train_ids <- data11_train$idauniq
test_ids <- data11_test$idauniq

stopifnot(nrow(data11_train) == length(split_train_ids))
stopifnot(nrow(data11_test) == length(split_test_ids))
stopifnot(identical(data11_train$idauniq, split_train_ids))
stopifnot(identical(data11_test$idauniq, split_test_ids))
stopifnot(length(intersect(data11_train$idauniq, data11_test$idauniq)) == 0)
```

# 1. Prepare test-set inputs

```{r}
standardise_qol_outcome <- function(data) {
  data %>%
    mutate(
      qol_status = case_when(
        as.character(qol_status) %in% c("Low QoL", "Low_QoL") ~ "Low_QoL",
        as.character(qol_status) %in% c("Non-low QoL", "Non_low_QoL") ~ "Non_low_QoL",
        TRUE ~ as.character(qol_status)
      ),
      qol_status = factor(qol_status, levels = c("Low_QoL", "Non_low_QoL")),
      across(where(is.ordered), ~ factor(.x, ordered = FALSE))
    )
}
```

```{r}
data11_train_cls <- data11_train %>%
  select(
    qol_status,
    all_of(full_feature_names)
  ) %>%
  standardise_qol_outcome()

data11_test_cls <- data11_test %>%
  select(
    qol_status,
    all_of(full_feature_names)
  ) %>%
  standardise_qol_outcome()

data11_selected_train <- data11_train %>%
  select(
    casp19_total,
    qol_status,
    all_of(selected_main_features)
  )

data11_selected_test <- data11_test %>%
  select(
    casp19_total,
    qol_status,
    all_of(selected_main_features)
  )

data11_selected_train_cls <- data11_selected_train %>%
  standardise_qol_outcome() %>%
  select(
    -casp19_total
  )

data11_selected_test_cls <- data11_selected_test %>%
  standardise_qol_outcome() %>%
  select(
    -casp19_total
  )
```

# 2. Evaluation helpers

```{r}
make_class_from_probability <- function(prob, threshold = 0.50) {
  factor(
    if_else(
      prob >= threshold,
      "Low_QoL",
      "Non_low_QoL"
    ),
    levels = c("Low_QoL", "Non_low_QoL")
  )
}

calculate_classification_metrics <- function(pred_df, model_label) {
  bind_rows(
    pred_df %>%
      roc_auc(truth = qol_status, .pred_Low_QoL, event_level = "first"),
    pred_df %>%
      pr_auc(truth = qol_status, .pred_Low_QoL, event_level = "first"),
    pred_df %>%
      accuracy(truth = qol_status, estimate = .pred_class),
    pred_df %>%
      precision(truth = qol_status, estimate = .pred_class, event_level = "first"),
    pred_df %>%
      sens(truth = qol_status, estimate = .pred_class, event_level = "first"),
    pred_df %>%
      spec(truth = qol_status, estimate = .pred_class, event_level = "first"),
    pred_df %>%
      f_meas(truth = qol_status, estimate = .pred_class, event_level = "first")
  ) %>%
    mutate(model = model_label) %>%
    select(model, .metric, .estimator, .estimate)
}

evaluate_fixed_workflow <- function(
    fitted_model,
    test_data,
    test_ids,
    model_id,
    model_label,
    threshold = 0.50
) {
  prob_df <- predict(
    fitted_model,
    new_data = test_data,
    type = "prob"
  )
  
  pred_df <- tibble(idauniq = test_ids) %>%
    bind_cols(prob_df) %>%
    bind_cols(test_data %>% select(qol_status)) %>%
    mutate(
      .pred_class = make_class_from_probability(
        .pred_Low_QoL,
        threshold = threshold
      ),
      model_id = model_id,
      model = model_label
    ) %>%
    select(
      idauniq,
      qol_status,
      model_id,
      model,
      .pred_Low_QoL,
      .pred_Non_low_QoL,
      .pred_class
    )
  
  list(
    predictions = pred_df,
    metrics = calculate_classification_metrics(pred_df, model_label),
    confusion = pred_df %>%
      count(model_id, model, qol_status, .pred_class, name = "n")
  )
}
```

```{r}
prepare_catboost_data <- function(train_data, test_data, outcome = "qol_status") {
  train_data <- train_data %>%
    mutate(
      qol_status = case_when(
        as.character(qol_status) %in% c("Low QoL", "Low_QoL") ~ "Low_QoL",
        as.character(qol_status) %in% c("Non-low QoL", "Non_low_QoL") ~ "Non_low_QoL",
        TRUE ~ as.character(qol_status)
      ),
      qol_status = factor(qol_status, levels = c("Low_QoL", "Non_low_QoL"))
    )
  
  test_data <- test_data %>%
    mutate(
      qol_status = case_when(
        as.character(qol_status) %in% c("Low QoL", "Low_QoL") ~ "Low_QoL",
        as.character(qol_status) %in% c("Non-low QoL", "Non_low_QoL") ~ "Non_low_QoL",
        TRUE ~ as.character(qol_status)
      ),
      qol_status = factor(qol_status, levels = c("Low_QoL", "Non_low_QoL"))
    )
  
  predictor_names <- setdiff(names(train_data), outcome)
  numeric_predictors <- predictor_names[map_lgl(train_data[predictor_names], is.numeric)]
  categorical_predictors <- setdiff(predictor_names, numeric_predictors)
  
  for (v in numeric_predictors) {
    med_v <- median(train_data[[v]], na.rm = TRUE)
    train_data[[v]][is.na(train_data[[v]])] <- med_v
    test_data[[v]][is.na(test_data[[v]])] <- med_v
  }
  
  for (v in categorical_predictors) {
    train_data[[v]] <- as.factor(train_data[[v]])
    test_data[[v]] <- as.factor(test_data[[v]])
    
    train_data[[v]] <- fct_na_value_to_level(train_data[[v]], level = "Missing")
    test_data[[v]] <- fct_na_value_to_level(test_data[[v]], level = "Missing")
    
    train_levels <- levels(train_data[[v]])
    
    if (!("New" %in% train_levels)) {
      train_levels <- c(train_levels, "New")
    }
    
    test_chr <- as.character(test_data[[v]])
    test_chr[!(test_chr %in% train_levels)] <- "New"
    test_data[[v]] <- factor(test_chr, levels = train_levels)
    train_data[[v]] <- factor(as.character(train_data[[v]]), levels = train_levels)
  }
  
  list(
    train = train_data,
    test = test_data
  )
}

get_catboost_probability <- function(model, pool) {
  prob <- catboost.predict(
    model,
    pool,
    prediction_type = "Probability"
  )
  
  as.numeric(prob)
}

evaluate_fixed_catboost <- function(
    fitted_model,
    train_raw,
    test_raw,
    test_ids,
    model_id,
    model_label,
    threshold = 0.50
) {
  cat_data <- prepare_catboost_data(
    train_data = train_raw,
    test_data = test_raw,
    outcome = "qol_status"
  )
  
  cat_test <- cat_data$test
  cat_test_pool <- catboost.load_pool(
    data = cat_test %>% select(-qol_status)
  )
  
  pred_prob <- get_catboost_probability(
    fitted_model,
    cat_test_pool
  )
  
  pred_df <- tibble(
    idauniq = test_ids,
    qol_status = cat_test$qol_status,
    model_id = model_id,
    model = model_label,
    .pred_Low_QoL = pred_prob,
    .pred_Non_low_QoL = 1 - pred_prob
  ) %>%
    mutate(
      .pred_class = make_class_from_probability(
        .pred_Low_QoL,
        threshold = threshold
      )
    )
  
  list(
    predictions = pred_df,
    metrics = calculate_classification_metrics(pred_df, model_label),
    confusion = pred_df %>%
      count(model_id, model, qol_status, .pred_class, name = "n")
  )
}
```

# 3. Evaluate fixed classification models

```{r}
cat_full_train_raw <- data11_train %>%
  select(
    qol_status,
    all_of(full_feature_names)
  ) %>%
  mutate(
    across(where(is.ordered), ~ factor(.x, ordered = FALSE))
  )

cat_full_test_raw <- data11_test %>%
  select(
    qol_status,
    all_of(full_feature_names)
  ) %>%
  mutate(
    across(where(is.ordered), ~ factor(.x, ordered = FALSE))
  )

cat_selected_train_raw <- data11_selected_train %>%
  mutate(across(where(is.ordered), ~ factor(.x, ordered = FALSE))) %>%
  select(
    -casp19_total
  )

cat_selected_test_raw <- data11_selected_test %>%
  mutate(across(where(is.ordered), ~ factor(.x, ordered = FALSE))) %>%
  select(
    -casp19_total
  )

stopifnot(
  file.exists(model_bundle$full_feature$catboost_model_file),
  file.exists(model_bundle$important_feature$catboost_model_file)
)

cat_full_model <- catboost.load_model(
  model_bundle$full_feature$catboost_model_file
)

cat_selected_model <- catboost.load_model(
  model_bundle$important_feature$catboost_model_file
)
```

```{r}
full_feature_evaluations <- list(
  lasso = evaluate_fixed_workflow(
    fitted_model = model_bundle$full_feature$lasso,
    test_data = data11_test_cls,
    test_ids = test_ids,
    model_id = "lasso_logistic_full",
    model_label = "Full-feature LASSO logistic regression (42)"
  ),
  random_forest = evaluate_fixed_workflow(
    fitted_model = model_bundle$full_feature$random_forest,
    test_data = data11_test_cls,
    test_ids = test_ids,
    model_id = "rf_full",
    model_label = "Full-feature Random Forest (42)"
  ),
  xgboost = evaluate_fixed_workflow(
    fitted_model = model_bundle$full_feature$xgboost,
    test_data = data11_test_cls,
    test_ids = test_ids,
    model_id = "xgb_full",
    model_label = "Full-feature XGBoost (42)"
  ),
  catboost = evaluate_fixed_catboost(
    fitted_model = cat_full_model,
    train_raw = cat_full_train_raw,
    test_raw = cat_full_test_raw,
    test_ids = test_ids,
    model_id = "catboost_full",
    model_label = "Full-feature CatBoost (42)"
  ),
  lightgbm = evaluate_fixed_workflow(
    fitted_model = model_bundle$full_feature$lightgbm,
    test_data = data11_test_cls,
    test_ids = test_ids,
    model_id = "lightgbm_full",
    model_label = "Full-feature LightGBM (42)"
  ),
  mlp = evaluate_fixed_workflow(
    fitted_model = model_bundle$full_feature$mlp,
    test_data = data11_test_cls,
    test_ids = test_ids,
    model_id = "mlp_full",
    model_label = "Full-feature MLP (42)"
  )
)
```

```{r}
important_feature_evaluations <- list(
  lasso = evaluate_fixed_workflow(
    fitted_model = model_bundle$important_feature$lasso,
    test_data = data11_selected_test_cls,
    test_ids = test_ids,
    model_id = "lasso_logistic_selected",
    model_label = "Important-feature LASSO logistic regression (15)"
  ),
  random_forest = evaluate_fixed_workflow(
    fitted_model = model_bundle$important_feature$random_forest,
    test_data = data11_selected_test_cls,
    test_ids = test_ids,
    model_id = "rf_selected",
    model_label = "Important-feature Random Forest (15)"
  ),
  xgboost = evaluate_fixed_workflow(
    fitted_model = model_bundle$important_feature$xgboost,
    test_data = data11_selected_test_cls,
    test_ids = test_ids,
    model_id = "xgb_selected",
    model_label = "Important-feature XGBoost (15)"
  ),
  catboost = evaluate_fixed_catboost(
    fitted_model = cat_selected_model,
    train_raw = cat_selected_train_raw,
    test_raw = cat_selected_test_raw,
    test_ids = test_ids,
    model_id = "catboost_selected",
    model_label = "Important-feature CatBoost (15)"
  ),
  lightgbm = evaluate_fixed_workflow(
    fitted_model = model_bundle$important_feature$lightgbm,
    test_data = data11_selected_test_cls,
    test_ids = test_ids,
    model_id = "lightgbm_selected",
    model_label = "Important-feature LightGBM (15)"
  ),
  mlp = evaluate_fixed_workflow(
    fitted_model = model_bundle$important_feature$mlp,
    test_data = data11_selected_test_cls,
    test_ids = test_ids,
    model_id = "mlp_selected",
    model_label = "Important-feature MLP (15)"
  )
)
```

```{r}
all_classification_evaluations <- c(
  full_feature_evaluations,
  important_feature_evaluations
)

final_classification_results_long <- map_dfr(
  all_classification_evaluations,
  "metrics"
)

final_classification_results_wide <- final_classification_results_long %>%
  mutate(
    .metric = case_when(
      .metric == "roc_auc" ~ "ROC_AUC",
      .metric == "pr_auc" ~ "PR_AUC",
      .metric == "accuracy" ~ "Accuracy",
      .metric == "precision" ~ "Precision",
      .metric == "sens" ~ "Sensitivity/Recall",
      .metric == "spec" ~ "Specificity",
      .metric == "f_meas" ~ "F1",
      TRUE ~ .metric
    ),
    .estimate = round(.estimate, 3)
  ) %>%
  select(model, .metric, .estimate) %>%
  pivot_wider(
    names_from = .metric,
    values_from = .estimate
  )

final_classification_results_wide
```

```{r}
classification_confusion_tables <- map_dfr(
  all_classification_evaluations,
  "confusion"
)

classification_test_predictions_all12 <- map_dfr(
  all_classification_evaluations,
  "predictions"
)

important_model_test_predictions_all6 <- map_dfr(
  important_feature_evaluations,
  "predictions"
)

important_prediction_coverage <- important_model_test_predictions_all6 %>%
  count(idauniq)

all_model_prediction_coverage <- classification_test_predictions_all12 %>%
  count(idauniq)

stopifnot(
  n_distinct(important_model_test_predictions_all6$model_id) == 6,
  nrow(important_prediction_coverage) == length(test_ids),
  all(important_prediction_coverage$n == 6),
  n_distinct(classification_test_predictions_all12$model_id) == 12,
  nrow(all_model_prediction_coverage) == length(test_ids),
  all(all_model_prediction_coverage$n == 12),
  !anyNA(classification_test_predictions_all12$.pred_Low_QoL),
  all(
    classification_test_predictions_all12$.pred_Low_QoL >= 0 &
      classification_test_predictions_all12$.pred_Low_QoL <= 1
  ),
  all(
    abs(
      classification_test_predictions_all12$.pred_Low_QoL +
        classification_test_predictions_all12$.pred_Non_low_QoL -
        1
    ) < 1e-6
  )
)
```

# 4. Continuous LASSO test-set evaluation

```{r}
data11_test_lm <- data11_test %>%
  select(
    casp19_total,
    all_of(full_feature_names)
  ) %>%
  mutate(
    across(where(is.ordered), ~ factor(.x, ordered = FALSE))
  )

continuous_lasso_test_predictions <- predict(
  model_bundle$continuous_lasso,
  new_data = data11_test_lm
) %>%
  bind_cols(
    data11_test_lm %>% select(casp19_total)
  )

continuous_lasso_test_metrics <- bind_rows(
  continuous_lasso_test_predictions %>%
    rmse(truth = casp19_total, estimate = .pred),
  continuous_lasso_test_predictions %>%
    mae(truth = casp19_total, estimate = .pred),
  continuous_lasso_test_predictions %>%
    rsq(truth = casp19_total, estimate = .pred)
) %>%
  mutate(model = "Full-feature LASSO linear regression") %>%
  select(model, .metric, .estimator, .estimate)

continuous_lasso_test_metrics
```

# 5. Export held-out test results

```{r}
saveRDS(
  classification_test_predictions_all12,
  file.path(processed_path, "09_model_test_predictions_all12.rds")
)

saveRDS(
  important_model_test_predictions_all6,
  file.path(processed_path, "09_important_model_test_predictions_all6.rds")
)

write_xlsx(
  list(
    classification_performance = final_classification_results_wide,
    classification_performance_long = final_classification_results_long,
    confusion_tables = classification_confusion_tables,
    continuous_lasso_metrics = continuous_lasso_test_metrics
  ),
  file.path(output_path, "09_heldout_test_model_performance.xlsx")
)
```
