---
title: "10_exploratory_lasso_weight_threshold"
author: "Yiwen Xiang"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: true
---

This file conducts an exploratory threshold and class-weight sensitivity analysis for the important-feature LASSO logistic regression model. It uses the same analytical dataset, train-test split, and 15-feature predictor set as `08_model_development_training_only.Rmd`, then compares alternative class weights and classification thresholds for the Low_QoL positive class.

The purpose is not to re-select features or compare all models again. The model type is read from the training-set CV ranking saved by script 08. The class weight, penalty, and classification threshold are selected using training-set cross-validation only. The resulting operating point is fixed before it is applied to the test set in this exploratory analysis.

# 1. Setup

```{r, include=FALSE}
source(here::here("script", "00_packages.R"))

set.seed(123)
options(yardstick.event_first = TRUE)
```

```{r}
data11_training <- readRDS(
  file.path(processed_path, "07_wave11_ml_data.rds")
)

main_features <- readRDS(
  file.path(processed_path, "08_main_features.rds")
)

stopifnot(
  length(main_features) == 15L,
  !anyDuplicated(main_features),
  all(main_features %in% names(data11_training))
)

model_selection <- readRDS(
  file.path(processed_path, "08_exploratory_model_selected_by_cv.rds")
)

selected_exploratory_model <- model_selection$selected_model

if (selected_exploratory_model$model_id[[1]] != "lasso_logistic_selected") {
  stop(
    paste0(
      "The training-CV selected model is ",
      selected_exploratory_model$model[[1]],
      ", not LASSO logistic regression."
    )
  )
}

dim(data11_training)
main_features
selected_exploratory_model
```

# 2. Use the same train-test split as script 08

The split is kept consistent with script 08 so the LASSO sensitivity results are comparable with the main model results. The saved split IDs from script 08 are reused directly.

```{r}
split_id_file <- file.path(processed_path, "08_train_test_split_ids.rds")

if (!file.exists(split_id_file)) {
  stop(
    paste(
      "The train-test split IDs from script 08 were not found.",
      "Run script 08 before script 10."
    )
  )
}

split_ids <- readRDS(split_id_file)

train_ids <- split_ids$train_idauniq
test_ids <- split_ids$test_idauniq

stopifnot(!anyDuplicated(data11_training$idauniq))

train_position <- match(
  train_ids,
  data11_training$idauniq
)

test_position <- match(
  test_ids,
  data11_training$idauniq
)

if (anyNA(train_position) || anyNA(test_position)) {
  stop("Some split IDs were not found in the modelling dataset.")
}

data11_train <- data11_training[
  train_position,
  ,
  drop = FALSE
]

data11_test <- data11_training[
  test_position,
  ,
  drop = FALSE
]

stopifnot(
  identical(data11_train$idauniq, train_ids),
  identical(data11_test$idauniq, test_ids)
)

dim(data11_train)
dim(data11_test)
```

```{r}
data11_train %>%
  count(qol_status) %>%
  mutate(percent = round(n / sum(n) * 100, 2))

data11_test %>%
  count(qol_status) %>%
  mutate(percent = round(n / sum(n) * 100, 2))
```

# 3. Helper functions

## 3.1 Standardise outcome

`standardise_qol_outcome()` keeps the outcome coding consistent across scripts. `Low_QoL` is set as the first factor level so that ROC-AUC, PR-AUC, sensitivity, precision, and F1 treat it as the positive class.

```{r}
standardise_qol_outcome <- function(data) {
  data %>%
    mutate(
      qol_status = case_when(
        as.character(qol_status) %in% c("Low QoL", "Low_QoL") ~ "Low_QoL",
        as.character(qol_status) %in% c("Non-low QoL", "Non_low_QoL") ~ "Non_low_QoL",
        TRUE ~ as.character(qol_status)
      ),
      qol_status = factor(qol_status, levels = c("Low_QoL", "Non_low_QoL")),
      across(where(is.ordered), ~ factor(.x, ordered = FALSE))
    )
}
```

## 3.2 Evaluate one threshold on the test set

`evaluate_binary_predictions()` calculates performance after converting predicted Low_QoL probabilities into classes using one fixed threshold.

```{r}
evaluate_binary_predictions <- function(pred_df, model_name, strategy, weight, threshold) {
  pred_df <- pred_df %>%
    mutate(
      .pred_class = factor(
        if_else(.pred_Low_QoL >= threshold, "Low_QoL", "Non_low_QoL"),
        levels = c("Low_QoL", "Non_low_QoL")
      )
    )
  
  bind_rows(
    pred_df %>%
      roc_auc(truth = qol_status, .pred_Low_QoL, event_level = "first"),
    
    pred_df %>%
      pr_auc(truth = qol_status, .pred_Low_QoL, event_level = "first"),
    
    pred_df %>%
      accuracy(truth = qol_status, estimate = .pred_class),
    
    pred_df %>%
      precision(truth = qol_status, estimate = .pred_class, event_level = "first"),
    
    pred_df %>%
      sens(truth = qol_status, estimate = .pred_class, event_level = "first"),
    
    pred_df %>%
      spec(truth = qol_status, estimate = .pred_class, event_level = "first"),
    
    pred_df %>%
      f_meas(truth = qol_status, estimate = .pred_class, event_level = "first")
  ) %>%
    mutate(
      model = model_name,
      strategy = strategy,
      class_weight = weight,
      threshold = round(threshold, 3),
      .estimate = round(.estimate, 3)
    ) %>%
    select(model, strategy, class_weight, threshold, .metric, .estimator, .estimate)
}
```

## 3.3 Evaluate thresholds from CV predictions

`make_threshold_table()` evaluates a grid of thresholds using cross-validated training predictions. The test set is not used to choose thresholds.

```{r}
make_threshold_table <- function(pred_df) {
  threshold_grid <- seq(0.05, 0.95, by = 0.01)
  
  map_dfr(
    threshold_grid,
    function(threshold_i) {
      class_df <- pred_df %>%
        mutate(
          .pred_class = factor(
            if_else(
              .pred_Low_QoL >= threshold_i,
              "Low_QoL",
              "Non_low_QoL"
            ),
            levels = c("Low_QoL", "Non_low_QoL")
          )
        )
      
      tibble(
        threshold = threshold_i,
        sensitivity = sens(
          class_df,
          truth = qol_status,
          estimate = .pred_class,
          event_level = "first"
        )$.estimate,
        specificity = spec(
          class_df,
          truth = qol_status,
          estimate = .pred_class,
          event_level = "first"
        )$.estimate,
        precision = precision(
          class_df,
          truth = qol_status,
          estimate = .pred_class,
          event_level = "first"
        )$.estimate,
        f1 = f_meas(
          class_df,
          truth = qol_status,
          estimate = .pred_class,
          event_level = "first"
        )$.estimate
      )
    }
  )
}
```

# 4. Prepare the 15-feature LASSO data

This script uses the 15 main features saved by script 08, so the model checked here is the important-feature LASSO logistic regression model.

```{r}
selected_main_features <- main_features

data11_selected <- data11_training %>%
  select(
    idauniq,
    casp19_total,
    qol_status,
    all_of(selected_main_features)
  )

data11_selected_train_cls <- data11_selected %>%
  filter(idauniq %in% data11_train$idauniq) %>%
  standardise_qol_outcome() %>%
  select(
    -idauniq,
    -casp19_total
  )

data11_selected_test_cls <- data11_selected %>%
  filter(idauniq %in% data11_test$idauniq) %>%
  standardise_qol_outcome() %>%
  select(
    -idauniq,
    -casp19_total
  )

dim(data11_selected_train_cls)
dim(data11_selected_test_cls)
```

# 5. Class-weight candidates

Low_QoL case weights are applied as individual case weights. A weight of 1 represents no additional weighting. The prevalence-based weight is calculated as the ratio of Non_low_QoL to Low_QoL participants in the training set.

```{r}
train_class_counts <- data11_selected_train_cls %>%
  count(qol_status)

prevalence_weight <- train_class_counts %>%
  summarise(
    weight = n[qol_status == "Non_low_QoL"] / n[qol_status == "Low_QoL"]
  ) %>%
  pull(weight)

weight_grid <- sort(unique(c(1, 2, round(prevalence_weight, 2), 3, 4)))

train_class_counts
prevalence_weight
weight_grid
```

# 6. Run LASSO threshold and weight sensitivity checks

For each candidate class weight, `run_lasso_weight_experiment()` retunes the LASSO penalty using stratified 10-fold cross-validation and then estimates threshold performance from the out-of-fold training predictions.

```{r}
classification_metrics <- metric_set(
  roc_auc,
  pr_auc,
  accuracy,
  precision,
  sens,
  spec,
  f_meas
)

lasso_model <- logistic_reg(
  mode = "classification",
  penalty = tune(),
  mixture = 1
) %>%
  set_engine("glmnet")

lasso_grid <- grid_regular(
  penalty(range = c(-5, 0)),
  levels = 50
)

run_lasso_weight_experiment <- function(class_weight) {
  weighted_train_data <- data11_selected_train_cls %>%
    mutate(
      case_weight = hardhat::importance_weights(
        if_else(
          qol_status == "Low_QoL",
          class_weight,
          1
        )
      )
    )
  
  lasso_recipe_i <- recipe(qol_status ~ ., data = weighted_train_data) %>%
    step_impute_median(all_numeric_predictors()) %>%
    step_unknown(all_nominal_predictors(), new_level = "Missing") %>%
    step_novel(all_nominal_predictors(), new_level = "New") %>%
    step_dummy(all_nominal_predictors()) %>%
    step_zv(all_predictors()) %>%
    step_normalize(all_numeric_predictors())
  
  set.seed(123)
  
  weighted_cv_folds <- vfold_cv(
    weighted_train_data,
    v = 10,
    strata = qol_status
  )
  
  lasso_workflow <- workflow() %>%
    add_case_weights(case_weight) %>%
    add_recipe(lasso_recipe_i) %>%
    add_model(lasso_model)
  
  set.seed(123)
  
  lasso_tune <- tune_grid(
    lasso_workflow,
    resamples = weighted_cv_folds,
    grid = lasso_grid,
    metrics = classification_metrics,
    control = control_grid(save_pred = TRUE)
  )
  
  selected_penalty <- select_best(
    lasso_tune,
    metric = "roc_auc"
  )
  
  oof_predictions <- collect_predictions(
    lasso_tune,
    parameters = selected_penalty
  ) %>%
    select(qol_status, .pred_Low_QoL, .pred_Non_low_QoL)
  
  threshold_results <- make_threshold_table(oof_predictions)
  
  list(
    class_weight = class_weight,
    workflow = lasso_workflow,
    selected_penalty = selected_penalty,
    
    tuning_results = collect_metrics(lasso_tune) %>%
      mutate(model = "Important-feature LASSO logistic regression (15)", class_weight = class_weight),
    
    threshold_grid = threshold_results %>%
      mutate(
        model = "Important-feature LASSO logistic regression (15)",
        class_weight = class_weight,
        penalty = selected_penalty$penalty[[1]]
      )
  )
}
```

```{r}
set.seed(123)

lasso_cv_list <- map(
  weight_grid,
  run_lasso_weight_experiment
)

lasso_tuning_results <- map_dfr(lasso_cv_list, "tuning_results")
lasso_threshold_grid <- map_dfr(lasso_cv_list, "threshold_grid")

default_threshold_results_across_weights <- lasso_threshold_grid %>%
  filter(abs(threshold - 0.50) < 1e-8)

default_threshold_results_across_weights
```

# 7. Select one operating point using training CV

The class weight, penalty, and classification threshold are selected using training-set cross-validation only. The pre-specified rule is to retain thresholds with sensitivity of at least 0.70 in cross-validation, then select the combination with the highest F1-score. Precision and specificity are used only to break ties.

```{r}
eligible_operating_points <- lasso_threshold_grid %>%
  filter(
    !is.na(sensitivity),
    !is.na(f1),
    !is.na(precision),
    !is.na(specificity),
    sensitivity >= 0.70
  )

if (nrow(eligible_operating_points) == 0) {
  stop(
    "No class-weight and threshold combination achieved CV sensitivity >= 0.70."
  )
}

selected_operating_point <- eligible_operating_points %>%
  arrange(
    desc(f1),
    desc(precision),
    desc(specificity),
    class_weight,
    threshold
  ) %>%
  slice_head(n = 1) %>%
  mutate(
    strategy = "Training-CV selected: sensitivity >= 0.70 and max F1"
  )

selected_operating_point
```

```{r}
selected_weight <- selected_operating_point$class_weight[[1]]
selected_threshold <- selected_operating_point$threshold[[1]]
selected_penalty <- selected_operating_point$penalty[[1]]

weight_values <- map_dbl(
  lasso_cv_list,
  ~ .x$class_weight
)

selected_index <- which.min(
  abs(weight_values - selected_weight)
)

selected_result <- lasso_cv_list[[selected_index]]

stopifnot(
  isTRUE(
    all.equal(
      selected_penalty,
      selected_result$selected_penalty$penalty[[1]]
    )
  )
)

selected_weight
selected_threshold
selected_penalty
selected_result$selected_penalty
```

# 8. Final model fitting and exploratory test-set evaluation

```{r}
final_weighted_train_data <- data11_selected_train_cls %>%
  mutate(
    case_weight = hardhat::importance_weights(
      if_else(
        qol_status == "Low_QoL",
        selected_weight,
        1
      )
    )
  )

final_lasso_workflow <- finalize_workflow(
  selected_result$workflow,
  selected_result$selected_penalty
)

set.seed(123)

final_lasso_fit <- fit(
  final_lasso_workflow,
  data = final_weighted_train_data
)

final_test_prob <- predict(
  final_lasso_fit,
  new_data = data11_selected_test_cls,
  type = "prob"
) %>%
  bind_cols(
    data11_selected_test_cls %>%
      select(qol_status)
  )

final_test_metrics <- evaluate_binary_predictions(
  pred_df = final_test_prob,
  model_name = "Important-feature LASSO logistic regression (15)",
  strategy = selected_operating_point$strategy[[1]],
  weight = selected_weight,
  threshold = selected_threshold
)

final_test_metrics
```

# 9. Export results

```{r}
sensitivity_export_file <- file.path(
  output_path,
  "10_exploratory_lasso_weight_threshold_results.xlsx"
)

write_xlsx(
  list(
    selected_point_cv = selected_operating_point,
    final_test_metrics = final_test_metrics,
    default_threshold_cv = default_threshold_results_across_weights,
    cv_threshold_grid = lasso_threshold_grid,
    lasso_tuning = lasso_tuning_results
  ),
  sensitivity_export_file
)

sensitivity_export_file
```

# 10. Interpretation notes

This analysis is considered exploratory because the class-weight and threshold analysis was added after reviewing the initial held-out test-set model comparison in script 09. However, the model type was selected in script 08 using training-set cross-validation only, and the class weight, penalty, and classification threshold were also determined using training-set cross-validation only. The fixed operating point was subsequently applied to the same held-out test set used in script 09. Therefore, the resulting performance is interpreted as exploratory rather than independent confirmatory validation.
